### 2021-02-21 A1/article.tex
🎉 Add placeholder LaTeX file for A1.

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\newcommand{\p}{\partial}
\usepackage{etoolbox}
\usepackage{algorithm}
\usepackage{changepage}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}

\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
% \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\providecommand{\myceil}[1]{\left \lceil #1 \right \rceil }
\providecommand{\myfloor}[1]{\left \lfloor #1 \right \rfloor }

\begin{document}

\title{\textbf{CSDS 491: Assignment 1}}

\author{Shaochen (Henry) ZHONG, ilc{sxz517@case.edu}}

\date{Due and submitted on 02/22/2021 \\ Spring 2021, Dr. Lewicki}
\maketitle

\end{document}

### 2021-02-27 A1/code/q3.py
⚙️ Add code scripts for A1.

@@ -1,0 +1,32 @@
+p_BT = 0.6
+p_MT = 0.2
+
+p_K__BF_MF = 0.3
+p_K__BF_MT = 0.2
+p_K__BT_MF = 0.6
+p_K__BT_MT = 0.1
+
+def find_p_B__K(p_BT, p_MT, p_K__BF_MF, p_K__BF_MT, p_K__BT_MF, p_K__BT_MT):
+    top = p_BT * ((p_K__BT_MT) * (p_MT) + (p_K__BT_MF) * (1-p_MT))
+    bottom = top + (1-p_BT) * ((p_K__BF_MT) * (p_MT) + (p_K__BF_MF) * (1-p_MT))
+
+    return top/bottom
+
+print(find_p_B__K(p_BT, p_MT, p_K__BF_MF, p_K__BF_MT, p_K__BT_MF, p_K__BT_MT))
+
+new_p_K__BF_MF = 0.01
+new_p_K__BF_MT = 0.05
+print(find_p_B__K(p_BT, p_MT, new_p_K__BF_MF, new_p_K__BF_MT, p_K__BT_MF, p_K__BT_MT))
+
+
+def find_p_M__K(p_BT, p_MT, p_K__BF_MF, p_K__BF_MT, p_K__BT_MF, p_K__BT_MT):
+    top = p_MT * ((p_K__BT_MT) * (p_BT) + (p_K__BF_MT) * (1-p_BT))
+    bottom = top + (1-p_MT) * ((p_K__BT_MF) * (p_BT) + (p_K__BF_MF) * (1-p_BT))
+
+    return top/bottom
+
+print(find_p_M__K(p_BT, p_MT, p_K__BF_MF, p_K__BF_MT, p_K__BT_MF, p_K__BT_MT))
+print(find_p_M__K(p_BT, p_MT, new_p_K__BF_MF, new_p_K__BF_MT, p_K__BT_MF, p_K__BT_MT))
+
+print(find_p_M__K(p_BT, p_MT, new_p_K__BF_MF, p_K__BF_MT, p_K__BT_MF, p_K__BT_MT))
+print(find_p_M__K(p_BT, p_MT, p_K__BF_MF, new_p_K__BF_MT, p_K__BT_MF, p_K__BT_MT))

### 2021-02-27 A1/code/q4.py

@@ -1,0 +1,51 @@
+from scipy import stats
+import matplotlib.pyplot as plt
+import numpy as np
+
+def p_theta__beta(theta, y, n, alpha, beta):
+    return stats.beta.pdf(theta, alpha + y, beta + n - y)
+
+
+def plot_p_theta__beta(alpha, beta):
+    x = np.linspace(0, 1, 100)
+    y = stats.beta.pdf(x, alpha, beta)
+
+    plt.xlabel("theta")
+    plt.ylabel("p(theta)")
+    plt.title(f"alpha = {alpha}, beta = {beta}")
+    plt.plot(x, y)
+    plt.show()
+
+
+# plot_p_theta__beta(8, 2)
+# plot_p_theta__beta(2, 8)
+# plot_p_theta__beta(0.1, 0.1)
+# plot_p_theta__beta(5, 5)
+# plot_p_theta__beta(1, 1)
+
+
+trial_list = [0, 1, 2, 5, 10, 100]
+
+def plot_p_theta__beta_bernoulli(alpha, beta, theta, trial_list):
+    f = stats.bernoulli(theta)
+    data = f.rvs(1000)
+    x = np.linspace(0, 1, 100)
+
+    for i in range(len(trial_list)):
+        n = trial_list[i]
+        y = sum(data[:n])
+        posterior = p_theta__beta(x, y, n, alpha, beta)
+        plt.plot(x, posterior, label = f'n = {n}')
+    plt.legend()
+    plt.xlabel("p(theta | y, n)")
+    plt.ylabel("theta")
+    plt.title(f"alpha = {alpha}, beta = {beta}")
+
+    plt.show()
+
+# plot_p_theta__beta_bernoulli(2, 8, 0.5, trial_list)
+# plot_p_theta__beta_bernoulli(8, 2, 0.5, trial_list)
+# plot_p_theta__beta_bernoulli(0.1, 0.1, 0.5, trial_list)
+# plot_p_theta__beta_bernoulli(5, 5, 0.5, trial_list)
+# plot_p_theta__beta_bernoulli(1, 1, 0.5, trial_list)
+

### 2021-02-27 A1/code/q5.py

@@ -1,0 +1,33 @@
+from scipy import stats
+import matplotlib.pyplot as plt
+import numpy as np
+
+def plot_poisson(lam, x_max = 12):
+    x = np.arange(0, x_max)
+    y = stats.poisson.pmf(x, lam)
+    plt.title(f"lambda = {lam}")
+    plt.bar(x, y)
+    plt.show()
+
+
+
+
+# plot_poisson(2)
+# plot_poisson(3)
+# plot_poisson(4)
+
+e = [0.53, 0.65, 0.91, 1.19, 1.30, 1.33, 1.90, 2.01, 2.48]
+alpha, _, beta = stats.gamma.fit(e, floc=0)
+
+def plot_gamma(alpha, beta, T):
+    x = np.linspace(0, 9, 1000)
+    n = np.arange(10)
+
+    for t in T:
+        y = stats.gamma.pdf(n, alpha, scale = beta + t)
+        plt.plot(n, y, label = f"T = {t}")
+    plt.legend()
+    plt.show()
+
+
+plot_gamma(alpha, beta, [0, 0.5, 1.5])

### 2021-02-27 A1/article.tex
✨ Add answers for A1.

@@
 
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{frame=tb,
+  language=Python,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{blue},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true,
+  tabsize=3
+}
+
 \begin{document}
@@
 
+% We've received some common questions for A1.  Here are some hints to help you go in the right direction.
+%
+% Q2.2: You need only prove there is a logical gap that would require an additional assumption.
+%
+% Q4: It is helpful to use the conjugate prior as shown in lecture.  The  normalization factor for the posterior then has the same form as the prior (but with different arguments).
+%
+% Q5: Remember that the data consists of event times, and the posterior is in terms of the Poisson rate parameter lambda given the number of observed data events in a given duration from 0 to T.
+You only need a single Poisson likelihood combined with the prior to get the posterior.
+%
+% Exploration: The terms discrete and continuous refer to the variables in the model.
+See the rubric for the grading criteria.
+Try to make your inference problems simple models of real-world situations, rather than just numerical examples.
+
+\section*{Q1}
+
+\subsection*{1.1.}
+
+\begin{align*}
+    p(x, y  \mid  z) &= \frac{p(x, y, z)}{p(x)} \\
+    &= \frac{p(x, y, z) \cdot p(x, z)}{p(x) \cdot p(x, z)} \\
+    &= \frac{p(x, z)}{p(z)} \cdot \frac{p(x, y, z)}{p(x, z)} \\\
+    &= p(x  \mid  z) p(y  \mid  x, z)
+\end{align*}
+
+\subsection*{1.2.}
+
+\begin{align*}
+    p(x  \mid  y, z) &= \frac{p(x, y, z)}{p(y, z)} \\
+    &= \frac{\frac{p(x, y, z)}{p(z)}}{\frac{p(y, z)}{p(z)}} = \frac{\frac{p(x, y, z)}{p(x, z)}\frac{p(x, z)}{p(z)}}{p(y  \mid  z)} \\
+    &= \frac{p(y  \mid  x, z) p(x  \mid  z)}{p(y  \mid  z)}
+\end{align*}
+
+\section*{Q1}
+\subsection*{2.1.}
+
+Assume $a = \{1, 2, 3 \}$, $b = \{1, 4 \}$, $c = \{1, 2, 6 \}$ from a set of $\{1, 2, 3, 4, 5, 6\}$, we have:
+
+\begin{align*}
+    p(a, b) &= \frac{\{1\}}{\{1, 2, 3, 4, 5, 6\}} = \frac{1}{6} = p(a)p(b) = \frac{1}{2} \cdot \frac{1}{3} \Longrightarrow a \bot b \\
+    p(b, c) &= \frac{\{1\}}{\{1, 2, 3, 4, 5, 6\}} = \frac{1}{6} = p(b)p(c) = \frac{1}{2} \cdot \frac{1}{3} \Longrightarrow b \bot c \\
+    p(a, c) &= \frac{\{1, 2\}}{\{1, 2, 3, 4, 5, 6\}} = \frac{1}{3} \\
+    p(a)p(c) &= \frac{1}{2}\cdot \frac{1}{2} = \frac{1}{4} \\
+    \Longrightarrow p(a, c) &\neq p(a)p(c)
+\end{align*}
+
+This suggests $a \bot b \wedge b \bot c \not \Rightarrow a \bot c$. For an example, we may have
+
+\begin{itemize}
+    \item \textbf{a}: Rainning tomorrow.
+    \item \textbf{b}: Had pasta as dinner today.
+    \item \textbf{c}: Not rainning tomorrow.
+\end{itemize}
+
+By common sense we know that $a \bot b \wedge b \bot c$ holds true, but we can't have $a \bot c$ as it must be one way or another in terms rainning tomorrow or not.
+
+\subsection*{2.2.}
+
+Assume $a \bot b  \mid  c$, we have:
+\begin{align*}
+    p(a, (b  \mid  c)) &= p(a) p(b  \mid  c)\\
+    &= p(a)\frac{p(b, c)}{p(c)}
+\end{align*}
+
+To have $a \bot b$, we must show $p(a, b) = p(a)p(b)$. The only way to convert the above equation to such format is to assume $b \bot c$ so that we can have:
+
+\begin{align*}
+    p(a, (b  \mid  c)) &= p(a)\frac{p(b, c)}{p(c)} \\
+    p(a, b) &= p(a) \frac{p(b)p(c)}{p(c)} \\
+    p(a, b) &= p(a)p(b)
+\end{align*}
+
+But $b \bot c$ is an assumption which cannot be guaranteed.
+
+\section*{Q3}
+
+Please refer to \ilc{code/q3.py} for code.
+
+\subsection*{3.1.}
+
+\ilc{0.7281553398058251}
+
+\subsection*{3.2.}
+\ilc{0.9765625}
+
+I am aming to have $p(B  \mid  K) > 0.9$, thus I will need to lower the possibility of scenarios where the butler is not the killer.
+I found $p(K  \mid  B = F, M = F)$ and $p(K  \mid  B = F, M = T)$ to be more ``explainable'' as for the former we may simply say the inspector had more on-scence info indicating the killer is very likely to be still in the house; for the latter we may also just let the scence implies that the victim is strong and tall but the maid is petite with no sign of combat -- thus lowering the possibility of $M=T$ along.
+
+\subsection*{3.3.}
+
+\begin{align*}
+    p(M  \mid  K) &= \sum_{b}p(b,M \mid K) =\sum_{b}\frac{p(b,M,K)}{p(K)} \\
+    &= \frac{\sum_{b}p(K \mid b,M)p(b,M)}{\sum_{b,m}p(K \mid b,m)p(b,m)} = \frac{p(M)\sum_{b}p(K \mid b,M)p(b)}{\sum_{m}p(m)\sum_{b}p(K \mid b,m)p(b)}
+\end{align*}
+
+\subsection*{3.4.}
+
+With the setup proposed by the textbook, we have $p(M \mid K)$ being  \ilc{0.06796116504854369}; with the setup defined by me, we have $p(M \mid K)$ being \ilc{0.05208333333333333}.\newline
+
+The main contributing factor is $p(K  \mid  B = F, M = T)$, as we have lowered it in the second setup, we directly decrease the chances of the maid being the murder and thus potentially lowered $p(M \mid K)$.
+We may also confirm this guessing by calculating only with the lowered $p(K  \mid  B = F, M = F)$ and only with the lowered $p(K  \mid  B = F, M = T)$.
+We got \ilc{0.08771929824561403} with only the lowered $p(K  \mid  B = F, M = F)$ and \ilc{0.04} with only the lowered $p(K  \mid  B = F, M = T)$ -- suggesting that knowing the maid is very unlikely to be the murder herslef ($\downarrow p(K  \mid  B = F, M = T)$) has more of an impact on $p(M \mid K)$ than knowing the murder is more likely to be between bulter and maid ($\downarrow p(K  \mid  B = F, M = F)$).
+
+\section*{Q4}
+
+Please refer to \ilc{code/q4.py} for complete code.
+
+\subsection*{4.1.}
+
+\begin{lstlisting}
+    def p_theta__beta(theta, y, n, alpha, beta):
+        return stats.beta.pdf(theta, alpha + y, beta + n - y)
+\end{lstlisting}
+
+\subsection*{4.2.}
+
+\begin{figure}[H]
+\minipage{0.4\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.2_1}.png}
+  \caption*{Biased coin, more likely tail.}
+\endminipage\hfill
+\minipage{0.4\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.2_2}.png}
+  \caption*{Biased coin, more likely head.}
+\endminipage
+\end{figure}
+
+Alternatively we may have it on the same plot with $\alpha < 1, \beta < 1$:
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.4\linewidth]{{fig/q4.2_1_alt}.png}\
+    \caption*{Biased coin (either way).}
+\end{figure}
+
+\begin{figure}[H]
+\minipage{0.4\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.2_3}.png}
+  \caption*{Unbiased coin}
+\endminipage\hfill
+\minipage{0.4\textwidth}
+    \includegraphics[width=\linewidth]{{fig/q4.2_4}.png}
+    \caption*{No info coin}
+\endminipage
+\end{figure}
+
+\subsection*{4.3.}
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.3_1}.png}
+  \caption*{Biased coin, more likely tail.}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.3_2}.png}
+  \caption*{Biased coin, more likely head.}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.3_1_alt}.png}
+  \caption*{Biased coin (either way).}
+\endminipage
+\end{figure}
+
+\begin{figure}[H]
+\minipage{0.4\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q4.3_3}.png}
+  \caption*{Unbiased coin}
+\endminipage\hfill
+\minipage{0.4\textwidth}
+    \includegraphics[width=\linewidth]{{fig/q4.3_4}.png}
+    \caption*{No info coin}
+\endminipage
+\end{figure}
+
+\subsection*{4.4.}
+
+Yes, as $n$ getting large we may more info on the likelihood and therefore result in an more accurate estimation.
+We may confirm this by checking at $n = 100$, regardless which prior we used, the prediction came out to be unbiased and accurate (around $\theta = 0.5$).
+
+\section*{Q5}
+
+Please refer to \ilc{code/q5.py} for code.
+
+\subsection*{5.1.}
+
+Let $x$ to be the actual occurence of events per unit time, we have:
+
+\begin{equation*}
+    p(x \mid \lambda) = \frac{e^{-\lambda} \lambda^x}{x!}
+\end{equation*}
+
+Since we have $9$ events occured in a duration of $3$ seconds, we estinate $\lambda = 3$.
+I will also plut $\lambda = 2$ and $\lambda = 4$ as examples of ``less'' and ``greater than'' of my estimation.
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q5.1_1}.png}
+  \caption*{$\lambda = 2$}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q5.1_2}.png}
+  \caption*{$\lambda = 3$}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/q5.1_3}.png}
+  \caption*{$\lambda = 4$}
+\endminipage
+\end{figure}
+
+\subsection*{5.2.}
+
+\begin{align*}
+    p(\lambda \mid n, T, \alpha, \beta) &= \frac{p(n, T \mid \lambda, \alpha, \beta) p(\lambda \mid \alpha, \beta)p(\alpha, \beta)}{p(n,T \mid \alpha, \beta)p(\alpha, \beta)}\\
+    &= \frac{p(n, T \mid \lambda, \alpha, \beta) \Gamma (\lambda; \alpha, \beta)}{p(n, T \mid \alpha, \beta)}\\
+    &= \frac{p(n, T \mid \lambda, \alpha, \beta) \Gamma (\lambda; \alpha, \beta)}{\int_0^{\infty}p(n, T\mid \lambda, \alpha, \beta)p(\lambda \mid \alpha, \beta) \cdot d\lambda} \\
+    &= \frac{ \frac{e^{-\lambda T}(\lambda{T})^n }{n!} \cdot \Gamma(\lambda; \alpha,\beta)}{\int_0^{\infty}(\frac{e^{-\lambda{T}}(\lambda{T})^n}{n!}  \cdot \Gamma (\lambda ; \alpha,\beta))d\lambda} \\
+    &= \frac{ e^{-\lambda{T}}(\lambda T )^n \cdot \lambda^{\alpha-1} e^{-\beta \lambda}}{\int_0^{\infty} e^{-\lambda (\beta + T)}(T \lambda)^n  \lambda^{\alpha-1} \cdot d\lambda}
+\end{align*}
+
+\subsection*{5.3.}
+
+\begin{align*}
+    p(\lambda \mid n, T, \alpha, \beta) &= \frac{ e^{-\lambda{T}}(\lambda T )^n \cdot \lambda^{\alpha-1} e^{-\beta \lambda}}{\int_0^{\infty} e^{-\lambda (\beta + T)}(T \lambda)^n  \lambda^{\alpha-1} \cdot d\lambda} \\
+    &= \frac{\frac{(\beta + T)^{\alpha+n}T^n \lambda^{n}e^{-\lambda(\beta+T)}\lambda^{\alpha-1}}{\Gamma(a+n)}}{T^n \cdot \frac{\Gamma(a + n)}{\Gamma(a)}\frac{\beta^{\alpha}}{(T + \beta)^{a+n}}} \\
+    % {T^n \int_0^{\infty} \frac{(\beta+T)^{\alpha+n}\lambda^n e^{-\lambda(\beta+T)}}{\Gamma(a+n)}d\lambda} \\
+    &= \frac{(\beta+T)^{\alpha+n}\lambda^{\alpha+n-1}e^{-\lambda(\beta+T)}}{\Gamma(a+n)} \\ \\
+    &= \Gamma(\lambda; \alpha+n, \beta+T)
+\end{align*}
+
+Thus, the Gamma distribution is the conjugate prior of Possion distribution with $\alpha' = \alpha + n$ and $\beta' = \beta + T$.
+
+\subsection*{5.4.}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.6\linewidth]{{fig/q5.4}.png}
+\end{figure}
+
+Although no new event has occured, the fact that no event has occured bring in useful information to the estimation.
+
+\section*{Q6}
+
+\textit{The following example is altered from the book}.\newline
+
+\noindent Assuem we have two 3-side dice to roll where we the first dice being $X \in \{1, 2, 3\}$ and $Y$ being the sum of two dices. We have the joint probability being:
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{ c | c | c | c}
+        \hline
+        $p(x, y)$ & $X = 1$ & $X = 2$ & $X = 3$ \\
+        \hline
+        $Y = 2$ & 1/9 & 0 & 0 \\
+        $Y = 3$ & 1/9 & 1/9 & 0 \\
+        $Y = 4$ & 1/9 & 1/9 & 1/9 \\
+        $Y = 5$ & 0 & 1/9 & 1/9 \\
+        $Y = 6$ & 0 & 0 & 1/9 \\
+    \end{tabular}
+\end{table}
+
+I capture the structure in such a way that if $y - x > 3$, the probability will be $0$ as the max roll from second dice is only $3$.
+If $y - x \leq 3$, this means we can only have one possible second dice outcome to match with the first dice, so the join probability will bt $(\frac{1}{3})^2 = \frac{1}{9}$.
+
+For conditional probablity, we have $p(y | x) = \frac{p(x, y)}{x}$.
+But since given a certain $x$, we will have only one option, $(y - x)$, for the second dice; thus, we have $p(y | x) = \frac{1}{3}$ if such $(x, y)$ combination is at all possible.
+On the othe hand we have $p(x | y) = \frac{p(x, y)}{y}$.
+Since we only have one first-second dice combination given a possible $(x, y)$, $p(x | y) = p(x) \cdot p(\text{second dice outcome}) = \frac{1}{9}$.
+
+For marginal probablity for $P(x)$, is it universially $\frac{1}{3}$ due to the nature of being a fair dice. But for $P(Y)$ we have:
+
+\begin{equation*}
+    P(y) = \sum_{j} P(x_j, y)
+\end{equation*}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{ c | c | c | c | c | c}
+        \hline
+         & $Y = 2$ & $Y = 3$ & $Y = 4$ & $Y = 5$ & $Y = 6$ \\
+        \hline
+        $P(Y)$ & 1/9 & 2(1/9) & 3(1/9) & 2(1/9) & 1/9
+    \end{tabular}
+\end{table}
+
+The intuition behind this equation is to get a certain $Y$, you must add the all possible dice combination to form such $Y$ by chances.
+Which means we may simply add up each row of the joint probablity table to form a table for all $P(Y)$.
+
+\section*{Q7: Exploration}
+
+\subsection*{Discrete Inference Problem}
+
+I overhead the idea of ``draw a dice from a jar (of different dices) and roll it without checking the dice'' in classroom and thought it would be interesting.
+Although the proposed question in classroom was seemingly unsolvable, many interesting questions can be made base on this set up.\newline
+
+\noindent Assuming we have a jar of three dices, $\{D4, D6, D8 \}$ representing a 4-, 6-, 8-side fair dice respectively.
+We may draw a dice from the jar, roll it, record the outcome and put it back.
+Say we got an output sequence of
+
+\begin{equation*}
+    \{1, 6, 3\}
+\end{equation*}
+
+What is the most likely sequence of the dice rolled?\newline
+
+\noindent This question will be easy if the output sequence is $\{ 8, 8, 8\}$ so that we know the dice rolling sequence must be $\{D8, D8, D8 \}$ as no other dice can generate an outcome of $8$.
+This implies for $P(D \mid 8)$, we have $P_1(D4 \mid 8) = P(D6 \mid 8) = 0$ and $P(D8 \mid 8) = 1$
+
+Although we can definitively tell what will $P_1(D_{\{4, 6, 8\}} \mid 1)$ be, we may infer base on the structure of the dice:
+
+\begin{align*}
+    P_1(D4 \mid 1) &= \frac{1}{4} \\
+    P_1(D6 \mid 1) &= \frac{1}{6} \\
+    P_1(D8 \mid 1) &= \frac{1}{8}
+\end{align*}
+
+So we may safely conclude that for the first outcome of the sequence $\{1\}$, it is most likely generated by $D4$.
+Similarily, for $\{1, 6\}$, we may find the dice-rolling sequence that has the maximal possibility of generating this output sequence.
+We already know the first dice is most likely to be $D4$, then we have
+\begin{align*}
+    P_2((D4 \mid 1)) &= P(D4) \cdot P(1 \mid D4) \cdot P(D4 \to D4) \cdot  \cdot P(6 \mid D4) \\
+    P_2((D6 \mid 1)) &= P(D4) \cdot P(1 \mid D4) \cdot P(D4 \to D6) \cdot  \cdot P(6 \mid D6) \\
+    P_2((D8 \mid 1)) &= P(D4) \cdot P(1 \mid D4) \cdot P(D4 \to D8) \cdot  \cdot P(6 \mid D8) \\
+\end{align*}
+
+In this case, we have $P_2((D6 \mid 1))$ yieling the largest probability, so the dice-rolling sequence is mostly likely to be $\{D4, D6\}$...
+and using the same idea we will have an answer of $\{D4, D6, D4\}$.\newline
+
+This question is simple because we have a short output sequence to decode, and we may even simply brute force this problem by trying out every combinations.
+But the idea of calculating the most likely dice one by one can be extend to sequence of any length, and even with many more add on restructions like:
+
+\begin{itemize}
+    \item Not to use the same dice twice in a row: we may simply set $P(D_x \to D_x) = 0$.
+    \item Got mutiple dice of same structure in the jar: we may simply $P(D_x \to D_y) = \frac{\text{left over $Dy$ dices}}{\text{all dices - number of $Dx$ dices}}$.
+    \item There is uneven transition possibility between $D_x \to D_y$: we set $P(D_x \to D_x)$ accordingly.
+\end{itemize}
+
+This question will eventually become a dynamic programming problem to solve.
+
+\subsection*{Continuous Inference Problem}
+
+%
+% I came across some interesting educational resources of Gaussian naive bayes with multiple random variables, and I thought this might be a good place to reiterate what I learned.
+%
+% Assume we are trying to predict weather a person loves a certain movie base on the amount of popcorn, soda pop, and candy this person consumed.
+We will have a 3-dimension continuous inference problem as the amount of food consumption is continuous (we also assume each food consumption distribution is Normal distribution).
+%
+% For visualization, let color green represents the food consumption distribution of the group who loves this movie, and color red represents the  food consumption distribution of the group who does not.
+The \textbf{X} mark is the person to predict as he has consumed $C = \{20, 500, 25\}$ units of popcorn, soda, and candy respectively:
+%
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.6\linewidth]{{fig/q7}.png}
+% \end{figure}
+%
+% Intuitively, we would assume this person will belong to the group who love such movie as he has consumed a lor more popcorn and soda than the group who doesn't. To form the equation, we may have:
+%
+% \begin{align*}
+%     p(L \mid C) &= \frac{p(L) \cdot p(p = 20 \mid L) \cdot p(s = 500 \mid L) \cdot p(c = 25 \mid L)}{P(C)} \\
+%     &= \frac{p(L) \cdot p(p = 20 \mid L) \cdot p(s = 500 \mid L) \cdot p(c = 25 \mid L)}{\int p(C \mid L)p(L) \cdot dL} \\
+%     p(\hat{L} \mid C) &= \frac{p(\hat{L}) \cdot p(p = 20 \mid \hat{L}) \cdot p(s = 500 \mid \hat{L}) \cdot p(c = 25 \mid \hat{L})}{P(C)} \\
+%     &= \frac{p(\hat{L}) \cdot p(p = 20 \mid \hat{L}) \cdot p(s = 500 \mid \hat{L}) \cdot p(c = 25 \mid \hat{L})}{\int p(C \mid \hat{L})p(\hat{L}) \cdot d \hat{L}}
+% \end{align*}
+%
+% In this question we assumed $p(L) = p(\hat{L})$ as prior.
+
+I will try to discuss how gramma distribution is a conmjugate prior for exponential distribution.
+
+For $\theta \sim Gamma(\alpha, \beta)$ we have a prior of:
+
+\begin{equation*}
+    f_\Theta(\theta) = \frac{\beta^\alpha\theta^{\alpha - 1}e^{-\beta\theta}}{\Gamma(\alpha)}
+\end{equation*}
+
+It's likelihood is a exponential distribution:
+
+\begin{equation*}
+    f(X|\theta) = \theta^n e^{-\theta \sum x_i}
+\end{equation*}
+
+Now to find the posterior (likelihood times prior):
+
+\begin{align*}
+    f_{\Theta | X}(\theta | x) &\propto f_{x|\Theta}(x|\theta)\times f_\Theta(\theta) \\
+    f_{\Theta | X}(\theta | x) &\propto \theta^n e^{-\theta \sum x_i} \times \frac{\beta^\alpha\theta^{\alpha - 1}e^{-\beta\theta}}{\Gamma(\alpha)} \\
+    &= \theta^{n + \alpha - 1}e^{-\theta(\sum x_i + \beta)}
+\end{align*}
+
+We have a gamma distribution of $\Gamma( \alpha' = \alpha + n, \beta' = \beta + \sum x_i )$ being the conjugate prior.
+
 \end{document}

### 2021-03-08 A2/article.tex
✨ Add answers for A2 Q3 and Q4.

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\newcommand{\p}{\partial}
\usepackage{etoolbox}
\usepackage{algorithm}
\usepackage{changepage}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}

\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
% \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\providecommand{\myceil}[1]{\left \lceil #1 \right \rceil }
\providecommand{\myfloor}[1]{\left \lfloor #1 \right \rfloor }

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\begin{document}

\title{\textbf{CSDS 491: Assignment 2}}

\author{Shaochen (Henry) ZHONG, ilc{sxz517@case.edu}}

\date{Due and submitted on 03/08/2021 \\ Spring 2021, Dr. Lewicki}
\maketitle

% We've received some common questions for A1.  Here are some hints to help you go in the right direction.
%
% Q2.2: You need only prove there is a logical gap that would require an additional assumption.
%
% Q4: It is helpful to use the conjugate prior as shown in lecture.  The  normalization factor for the posterior then has the same form as the prior (but with different arguments).
%
% Q5: Remember that the data consists of event times, and the posterior is in terms of the Poisson rate parameter lambda given the number of observed data events in a given duration from 0 to T.  You only need a single Poisson likelihood combined with the prior to get the posterior.
%
% Exploration: The terms discrete and continuous refer to the variables in the model.  See the rubric for the grading criteria. Try to make your inference problems simple models of real-world situations, rather than just numerical examples.

\section*{Q1}

\section*{Q2}

\section*{Q3. Model Complexity, Free Parameters, and Simplifying Assumptions}

\subsection*{3.1.}

\begin{align*}
    p(x_1, x_2, \dots, x_N) &= p(x_1 \mid x_2, x_3, \dots, x_N) p(x_2, x_3, \dots, x_N) \\
    &= p(x_1 \mid x_2, x_3, \dots, x_N) p(x_2 \mid x_3, x_4, \dots, x_N) p(x_3, \dots, x_N) \\
    &= p(x_N) \cdot \prod_{i = 1}^{N-1} p(x_i \mid x_{i+1},  x_{i+2}, \dots, x_{N})
\end{align*}

\subsection*{3.2.}

This questions is essentially asking how many combination can we have for $\{ x_1, x_2, \dots, x_N \}$ given each $x_i$ has $K$ differrents states. As there will be $K^N$ combinations, there will be $K^N - 1$ parameters needed since the last state of the last node will not need another new parameter (it can be represent as the negation of all other states).

Similarily, we may continue upon the formula obtained in above question where $p(x_N) \cdot \prod_{i = 1}^{N-1} p(x_i \mid x_{i+1},  x_{i+2}, \dots, x_{N})$. Known that $x_N$ needs $(K-1)$ parameters to represent (since the last state can be represent as the negation of all other states) and a node with $i$ parents need $K^i$ parameters to represent.

\begin{align*}
    \sum_{i = 1}^{N} (K-1)(K^{N - i}) &= (K - 1)(K^{N - 1}) + (K - 1)(K^{N - 2}) + \dots + (K - 1)(K^0) \\
    &= (K-1) \cdot \frac{K^N - 1}{K - 1} = K^N - 1
\end{align*}

And $K^N - 1$ will be generalized as $O(K^N)$.

\subsection*{3.3.}

The $m$ root nodes each with $K$ stats will need $m(K - 1)$ parameters. The $N-m$ leaf nodes will each have $(K-1)$ parameters needed for itself and $K^m$ parameters needed for its $m$ of $K$-state parents. Thus, we will have:

\begin{equation*}
    m(K - 1 ) + (N-m)(K-1)K^m
\end{equation*}

parameters needed.

\subsection*{3.4.}

Since the formula of noisy-OR is given in a ``leaky'' fashion $p(x_i \mid \textrm{pa}({x_i})) = 1 - (1 - \mu_{i0}) \prod_{j \in \textrm{pa}(x_i)}(1 - \mu_{ij})^{x_j}$ where a $(1 - \mu_{i0})$ is involved. Each leaf nodes will have $(K-1)$ parameters needed for itself, and each of its parents will need $m^(k-1)$ parameters as each parent node has $k$ states (but since we only need a single parent at a time so we don't need another new parameter for the last state, so $k-1$). Last, we need one more parameter for the leak node $\mu_{i0}$ for each leaf node. All others remain unchanged and we have (note $K - 1 = 1$):

\begin{equation*}
    m(K-1) + (N-m)(K-1)(m^{K-1} + 1)= m + (N-m)(m + 1)
\end{equation*}

\section*{Q4. Models of Conditional Probability}

\subsection*{4.1.}

There must be at least a parent $x_j = 1$ as otherwise for $p(x_i = 1 \mid \textrm{pa}(x_i)) = 1 - (1 - \mu_{i0}) \cdot \prod_{j \in \textrm{pa}(x_i)} (1 - \mu_{ij})^0 = 1 - 1(1 - \mu_{i0})$ which will be very close to $0$; in fact, $p(x_i = 1 \mid \textrm{pa}(x_i))$ will be $0$ if we don't consider the leak node. So we know that there must be at least $x_j = 1$.\newline

This means if we have one $x_j = 1$ with a high $\mu_{ij} = \p(x_i = 1 \mid x_j = 1)$, we will have a very low $1 - \mu_{ij}$ and therefore bring the overall $p(x_i = 1 \mid \textrm{pa}(x_i))$ to be closer to zero. Since only one $x_i = x_j = 1$ with a high $\mu_{ij}$ will dramatically increase the overall output of  $p(x_i = 1 \mid \textrm{pa}(x_i))$, this works similar to the OR gate which will make $p(x_i = 1 \mid \textrm{pa}(x_i)) = 1$ with just one of its parent nodes being $1$. The noisy-OR is considered to be ``soft'' as it won't make the $p(x_i = 1 \mid \textrm{pa}(x_i))$ to be exactly 1 but only close to 1.

\subsection*{4.2.}

It is a leak node defined in the D\'{i}ez version of Leaky Noisy-OR function. It is aim to compensate the fact that in a complicate model, a $x_i$ can be 1 even if all of its parents are not 1 -- which can be roughly understand as a prior/bias base on prior knowlegde to the model.

Since we may assign a value to each $\mu_{i0}$, even if all parents of $x_i$ are zeros, we have:

\begin{align*}
    p(x_i \mid \textrm{pa}({x_i})) &= 1 - (1 - \mu_{i0}) \prod_{j \in \textrm{pa}(x_i)}(1 - \mu_{ij})^{x_j} \\
    &= 1 - (1 - \mu_{i0}) \prod_{j \in \textrm{pa}(x_i)}(1 - \mu_{ij})^0 \\
    &= \mu_{i0}
\end{align*}

Note it is only the D\'{i}ez version of leaky noisy-OR function as it was made on the assumption that the prior knowlegde will not have any effect on the output of modeled variables (the conditional output given the parents, namely all the $\mu_{ij}$s). Where Henrion considered the prior knowlegde will affect the output of all modeled variables as $p(x_i \mid \textrm{pa}({x_i})) = 1 - (1 - \mu_{i0}) \prod_{j \in \textrm{pa}(x_i)}(\frac{1 - \mu_{ij}}{1 - \mu_{i0}})^{x_j}$ where the $(1 - \mu_{i0})$ is part of every $u_{ij}$ outputs.

\subsection*{4.3.}

We may do the following transformation to the noisy-OR function and express it in a fashion of sum-of-weighted-inputs. For the ease of expression we will ignore the $1 - \mu_{i0}$ as it can be considered as a special parent and merge into to the $\prod_{j \in \{ \textrm{pa}(x_i) + x_0 \}}(1 - \mu_{ij})^{x_j}$. So we have the noisy-OR function being:

\begin{equation*}
    p(x_i \mid \textrm{pa}({x_i})) = 1 - \prod_{j \in \{ \textrm{pa}(x_i) + x_0 \}} (1 - \mu_{ij})^{x_j}
\end{equation*}

If we set $W_{ij} = - \ln(1 - \mu_{ij})$, we may express the nosiy-OR as (note for all $j$ we have $j \in \{ \textrm{pa}(x_i) + x_0 \}$):

\begin{align*}
    p(x_i \mid \textrm{pa}({x_i})) &= \rho (\sum_j W_{ij} \cdot x_{j}) \\
    \rho(z) &= 1 - e^{-z}
\end{align*}

We may confirm this is equivalent to the original noisy-OR function as:

\begin{align*}
    p(x_i \mid \textrm{pa}({x_i})) &= 1 - e^{-\sum_j W_{ij} \cdot x_{j}} \\
    &=1 - \frac{1}{e^{\sum_j -\ln(1 - \mu_{ij}) x_j}} \\
    &= 1 - \frac{1}{\prod_j \frac{1}{(1 - \mu_{ij}) ^ {x_j}}} \\
    &= 1 - \prod_j (1 - \mu_{ij})^{x_j}
\end{align*}

So essentially, we have the noisy-OR and sigmoid $\sigma(z) = \frac{1}{1 + e^{-z}} = 1 - \sigma(-z)$ both being a transformation function of a sum of weighted inputs. However, the sigmoid function is more general as is can take negative weights for $W_ij$, where for noisy-OR it can't since $W_{ij} = - \ln(1 - \mu_{ij})$. Moreover, noisy-OR can only take weights in the range of $(-\ln(0), -\ln(1)]$, where sigmoid can take any value to be its weights.

Due to the weight limitation issue, sigmoid can certainly compute unique functions where noisy-OR can't. But for model which needs extreme probablity (need $\{0, 1 \}$ as outputs), sigmoid function might not be applied due to it will need some extremely large negative/positive weights to bring the $\sigma(z)$ to 0 or 1. This can often be a problem if a model was operating base on the assumption of noisy-OR, and I will show one example in the below question where noisy-OR works but sigmoid doesn't; where the probablity output of the model is exactly like the prediction of noisy-OR.

\subsection*{4.4.}

First for a model which sigmoid works but noist-OR doesn't, the AND gate can be a perfect example where we have (where $x_{j1}, x_{j2}$ are parent nodes of $x_i$):

\begin{table}[H]
    \centering
    \begin{tabular}{ c  c | c }
        \hline
        $x_{j1}$ & $x_{j2}$ & $x_i$ \\
        \hline
        0 & 0 & 0 \\
        0 & 1 & 0 \\
        1 & 0 & 0 \\
        1 & 1 & 1
    \end{tabular}
\end{table}

To model this distribution with noisy-OR function, we have (note we ignore the leak node here to simplify the expression):

\begin{align*}
    p(x_i = 1 \mid x_{j1} = 0, x_{j2} = 0) &= 1 - (1 - p(x_i = 1 \mid x_{j1} = 0))^{0} \cdot (1 - p(x_i = 1 \mid x_{j2} = 0))^{0} = 0 \\
    p(x_i = 1 \mid x_{j1} = 0, x_{j2} = 1) &= 1 - (1 - p(x_i = 1 \mid x_{j1} = 0))^{0} \cdot (1 - p(x_i = 1 \mid x_{j2} = 1))^{1}\\ &= 1 - \frac{1}{2}= \frac{1}{2} & \text{Should be 0}\\
    p(x_i = 1 \mid x_{j1} = 1, x_{j2} = 0) &= 1 - (1 - p(x_i = 1 \mid x_{j1} = 1))^{1} \cdot (1 - p(x_i = 1 \mid x_{j2} = 0))^{0} \\ &= 1 - \frac{1}{2}= \frac{1}{2} & \text{Should be 0}\\
    p(x_i = 1 \mid x_{j1} = 1, x_{j2} = 1) &= 1 - (1 - p(x_i = 1 \mid x_{j1} = 1))^{1} \cdot (1 - p(x_i = 1 \mid x_{j2} = 1))^{1} \\ &= 1 - \frac{1}{2} \cdot \frac{1}{2}= \frac{3}{4} & \text{Should be 1}    % p(x_i = 1 \mid \textrm{pa}(x)) &= 1 - \frac{1}{2} \cdot \frac{1}{2}= \frac{3}{4} & \text{Should be $\frac{1}{4}$}
\end{align*}

As shown, the noisy-OR cannot model this system accurately, and this can't be solved by introducing a leak node -- since for the same $x_i$, the output answer has a different diffrence to the desired answer.\newline

\noindent However, if we use the sigmoid function to model this distribution, we may view the three neurons (events) as following:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{{fig/p4.4}.png}
\end{figure}

And we may set $W_{ij1} = W_{ij2} = m$ and set the bias of $x_i$ to be $-1.5m$; where such $m$ must be large enough such as $\sigma(0.5m) \approx 1$. Now to calculate the probablity outputs:

\begin{align*}
    p(x_i = 1 \mid x_{j1} = 0, x_{j2} = 0) &= \sigma(0 - 1.5m) \approx 0 \\
    p(x_i = 1 \mid x_{j1} = 0, x_{j2} = 1) &= \sigma(1m - 1.5m) \approx 0 \\
    p(x_i = 1 \mid x_{j1} = 1, x_{j2} = 0) &= \sigma(1m - 1.5m) \approx 0 \\
    p(x_i = 1 \mid x_{j1} = 1, x_{j2} = 1) &=  \sigma(2m - 1.5m) \approx 1
\end{align*}

Clearly sigmoid can model this distribution as desired. This is because in sigmoid we can take an arbitrarily large weight between neurons (in this case, $m$) but we can't do such thing in noisy-OR. \newline

\noindent However, there are also distributions that can only be nicely modeled by noisy-OR but not sigmoid. Intuitively, a distribution that follows noisy-OR must be modeable by the noist-OR function; and here's an example of how a distribution follows noisy-OR assumptions can't be modeled by sigmoid.

In the above section we already established that the noisy-OR function can also be expressed as sum-of-weighted-inputs:

\begin{align*}
    p(x_i \mid \textrm{pa}({x_i})) &= \rho (\sum_j W_{ij} \cdot x_{j}) \\
    \rho(z) &= 1 - e^{-z}
\end{align*}

So if we reuse the above diagram, but set $W_{ij1} = -\ln(1 - \mu{ij1})$ and $W_{ij2} = -\ln(1 - \mu{ij2})$, we may have a distribution of:

\begin{table}[H]
    \centering
    \begin{tabular}{ c  c | c }
        \hline
        $x_{j1}$ & $x_{j2}$ & $p(x_i)$ \\
        \hline
        0 & 0 & 0 \\
        0 & 1 & $1-e^{-1}$ \\
        1 & 0 & $1-e^{-1}$ \\
        1 & 1 & $1-e^{-2}$
    \end{tabular}
\end{table}

It is trivial to show that this distribution can be modeled by the noisy-OR function as it is defined with respect to noisy-OR. So the only thing left is to show that this distribution cannot be nicely modeled by the sigmoid function.

However, this is an impossible task because:

\begin{itemize}
    \item To have $p(x_i  \mid x_{j1} = 0, x_{j2} = 0) = 0$ we must have very large and negative sum-of-weighted-inputs $z$ to bring $\sigma(z - \text{}) \approx 0$. This can be done by setting a very large and negative bias on $x_i$, like $-m$.
    \item To have $p(x_i \mid x_{j1} = 0, x_{j2} = 1), p(x_i  \mid x_{j1} = 1, x_{j2} = 0) = 1-e^{-1}$, since $1-e^{-1}$ is a positive value, we must have the $W_{ij1}, W_{ij2}$ to be greater than the absolute value of the bias $|-m|$ so that we will have $\sigma(W_{ij1} - m)$ or $\sigma(W_{ij2} - m)$ to be $\geq 0$. This means $W_{ij1}, W_{ij2}$ must be some significantly large positive value that is $> m$.
    \item Since we have established that $W_{ij1}, W_{ij2}$ must be some significantly large positive value that is $> m$. For $p(x_i \mid x_{j1} = 1, x_{j2} = 1)$ with the sigmoid function we must have $\sigma(z = W_{ij1} + W_{ij2} - m)$; where such $z$ is guaranteed to be $>m$ and we know that $\sigma(m) \approx 1$. This will create an unavoidable confict with the purposed distribution as the output should be $1-e^{-2}$ which is $\ll 1$.
\end{itemize}

So we have showed an example distribution where the noisy-OR function works but the sigmoid function doesn't. This is because to have outputs like $\{0, 1\}$ sigmoid, we must relay on large weights and large bias; but since the distribution is defined with respect to noisy-OR, it was desiged to use with smaller weights.\newline

\noindent The second example was inspired by -- or technically, taken from (although I believe there a typo in the original text regarding $e^{-1}$ where it should be $1 - e^{-1}$) -- \href{http://www.cs.toronto.edu/~bonner/courses/2016s/csc321/readings/Connectionist%20learning%20of%20belief%20networks.pdf}{R.M. Neal, Connectionist learning of belief networks, Artificial Intelligence 56 (1992) 71–113.}

\section*{Q5}

\section*{Exploration}

\end{document}

### 2021-03-08 A2/article.tex
💄 Fix typo on email.
✨ Add answers for A2 Q1.

@@
 
-\author{Shaochen (Henry) ZHONG, ilc{sxz517@case.edu}}
+\author{Shaochen (Henry) ZHONG, \ilc{sxz517@case.edu}}
 
@@
 
-\section*{Q1}
+\section*{Q1. Conditional Independence}
 
-\section*{Q2}
+\subsection*{1.1.}
+
+\begin{align*}
+    p(a,b) &= \sum\limits_c \sum\limits_d \sum\limits_e p(a, b, c, d, e) \\
+    &= \sum\limits_c  p(c \mid a, b) p(a)p(b) \cdot \sum\limits_d p(d \mid c) \cdot \sum\limits_e p(e \mid c) \\
+    &\text{Known that} \ \sum\limits_{x} p(x \mid pa(x)) = 1 \\
+    &= p(a)p(b)
+\end{align*}
+
+Thus, $a \perp b \mid \emptyset$.
+
+\subsection*{1.2.}
+
+For $a \perp b \mid e$ we must have $p(a, b \mid e) = p(a \mid e) p(b \mid e)$. Which implies $p(a \mid b, e) = p(a \mid e)$ since:
+
+\begin{align*}
+    p(a \mid b) &= \frac{p(a, b)}{p(b)} \\
+    p(a \mid b, e) &= \frac{p(a, b \mid e)}{p(b \mid e)} \\
+    &= \frac{p(a \mid e) p(b \mid e)}{p(b \mid e)} \\
+    &= p(a \mid e)
+\end{align*}
+
+Known that $a \perp b \mid \emptyset$, now to calculate $p(a \mid b, e)$ and $p(a \mid e)$ to see if they are equal:
+
+\begin{align*}
+    p(a \mid b,e) &= \frac{p(a,b,e)}{p(b,e)} = \frac{\sum_c p(a,b,c,e)}{\sum_a\sum_c p(a,b,c,e)}\\
+    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_a\sum_c p(e \mid c)p(a,b,c)}\\
+    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_a\sum_c p(e \mid c)p(c \mid a, b) p(a, b)}\\
+    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_a\sum_c p(e \mid c)p(c \mid a, b) p(a) p(b)}\\
+    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_c p(e \mid c)p(b, c)}
+\end{align*}
+
+\begin{align*}
+    p(a \mid e) &= \frac{p(a,e)}{p(e)} = \frac{\sum_c\sum_b p(a,b,c,e)}{\sum_c p(e \mid c)p(c)}\\
+    &= \frac{\sum_c p(e \mid c) \sum_b p(a,b,c)}{\sum_c p(e \mid c)p(c)}\\
+    &= \frac{\sum_c p(e \mid c) p(a,c)}{\sum_c p(e \mid c)p(c)}
+\end{align*}
+
+Thus, it is possible that $p(a \mid b,e) \neq p(a \mid e)$ unless we may confirm that $p(b) = 1$.
+
+\section*{Q2. Conditional Independence and Causality}
+
+The conditional independence assumption this model made is $b \perp c \mid a$ as $p(b, c \mid a) = p(b \mid a) p(c \mid a)$; and joint distribution of this model is therefore $p(a, b, c) = p(b\mid  a)p(c\mid a)p(a)$.
+
+However, the same joint distribution $p(a, b, c)$ can also be modeled differently:
+
+\begin{align*}
+    p(a, b, c) &= p(a \mid b, c)p(b \mid c)p(c) \\
+    &= p(b \mid a, c)p(a \mid c)p(c) \\
+    &= p(c \mid a, b)p(a \mid b)p(b) \\
+    &= \dots
+\end{align*}
+
+Thus, the condition of the purposed graph does not necessarily hold.
 

### 2021-03-09 A2/code/q5.py
⚙️ Add code for A2 Q5.

@@ -1,0 +1,86 @@
+from pgmpy.models import BayesianModel
+from pgmpy.factors.discrete import TabularCPD
+from pgmpy.inference import VariableElimination
+
+model = BayesianModel([("Battery", "Gauge"),("Battery", "Turns Over"),("Turns Over", "Starts"), ("Fuel", "Gauge"),("Fuel", "Starts"), ])
+
+state_name_dict = {
+    'Battery': ['good', 'bad'],
+    'Fuel': ['not empty', 'empty'],
+    'Gauge': ['not empty', 'empty'],
+    'Turns Over': ['yes', 'no'],
+    'Starts': ['yes', 'no'],
+}
+
+
+cpd_b = TabularCPD(variable='Battery', variable_card=2, values=[[0.98], [0.02]], state_names={'Battery': state_name_dict['Battery']})
+
+# p(f = no e), p(f = e)
+cpd_f = TabularCPD(variable='Fuel', variable_card = 2, values=[[0.95], [0.05]], state_names={'Fuel': state_name_dict['Fuel']})
+
+# p(g | b, f)
+cpd_g = TabularCPD(variable='Gauge', variable_card=2,
+                    values=[[0.96, 0.03, 0.9, 0.01], [0.04, 0.97, 0.1, 0.99]],
+                    evidence=['Battery', 'Fuel'],
+                    evidence_card=[2, 2],
+                    state_names={'Gauge': state_name_dict['Gauge'],
+                                'Battery': state_name_dict['Battery'],
+                                'Fuel': state_name_dict['Fuel']}
+                                )
+# p(t | b)
+cpd_t = TabularCPD(variable='Turns Over', variable_card=2,
+                    values=[[0.97, 0.02], [0.03, 0.98]],
+                    evidence=['Battery'],
+                    evidence_card=[2],
+                    state_names={'Turns Over': state_name_dict['Turns Over'],
+                                'Battery': state_name_dict['Battery']}
+                                )
+
+# p(s)
+cpd_s = TabularCPD(variable='Starts', variable_card=2,
+                    values=[[0.99, 0.08, 0.0, 0.01], [0.01, 0.92, 1.0, 0.99]],
[…]

### 2021-03-09 A2/article.tex
✨ Add answers for A2 Q5.

@@
 
-\section*{Q5}
+\section*{Q5. Car Troubles}
+
+\subsection*{5.1.}
+
+\begin{align*}
+    p(f=\text{empty} \mid s=\text{no}) &= \frac{p(s=\text{no}, f=\text{empty})}{p(s=\text{no})}\\
+    &= \frac{\sum_t\sum_b p(s=\text{no}, f=\text{empty}, \ t, b)}{\sum_f\sum_t\sum_b p(s=\text{no},f,t,b)}\\
+    &= \frac{\sum_t\sum_b p(s=\text{no} \mid f=\text{empty},t)p(f, t)p(b)}{\sum_f\sum_t\sum_b p(s=\text{no} \mid f,t)p(f, t)p(b)}\\
+    &= \frac{\sum_t\sum_b p(s=\text{no} \mid f=\text{empty},t) \ p(f)p(t \mid b)p(b)}{\sum_f\sum_t\sum_b p(s=\text{no} \mid f,t) \ p(f)p(t \mid b)p(b)}\\
+    &= \frac{0.0461715}{0.101756} \approx 0.4537
+\end{align*}
+
+\subsection*{5.2.}
+
+The following result generated by \ilc{code/q5.py} confirms my calculation of the pervious question.
+
+\begin{lstlisting}
++-----------------+-------------+
+| Fuel            |   phi(Fuel) |
++=================+=============+
+| Fuel(not empty) |      0.5463 |
++-----------------+-------------+
+| Fuel(empty)     |      0.4537 |
++-----------------+-------------+
+\end{lstlisting}
+
+\subsection*{5.3.}
+
+Results within this question is generated by \ilc{code/q5.py}.
+
+\subsubsection*{Scenario 1: No fuel, given Turns Over first.}
+
+Given that \ilc{Turns Over = no} and known that \ilc{Starts = no}, we may query on the probability of \ilc{Battery} and \ilc{Fuel}:
+
+\begin{lstlisting}
++---------------+----------------+          +-----------------+-------------+
+| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
++===============+================+          +=================+=============+
+| Battery(good) |         0.6000 |          | Fuel(not empty) |      0.9505 |
++---------------+----------------+          +-----------------+-------------+
+| Battery(bad)  |         0.4000 |          | Fuel(empty)     |      0.0495 |
++---------------+----------------+         +-----------------+-------------+
+\end{lstlisting}
+
+So the problem is mostly likely caused by the car having no fuel.
+By asking our friend to provide additional information of \ilc{Gauge = not empty}, we may again query on the probability of \ilc{Battery} and \ilc{Fuel}:
+\begin{lstlisting}
++---------------+----------------+          +-----------------+-------------+
+| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
++===============+================+          +=================+=============+
+| Battery(good) |         0.6156 |          | Fuel(not empty) |      0.9988 |
++---------------+----------------+          +-----------------+-------------+
+| Battery(bad)  |         0.3844 |          | Fuel(empty)     |      0.0012 |
++---------------+----------------+          +-----------------+-------------+
+\end{lstlisting}
+
+We may notice that our confidence of the car having no fuel has further improved, and we have therefore found the true cause of the problem.
+
+\subsubsection*{Scenario 2: Bad Battery, given Gauge first.}
+
+Similar to the above setup but this time we were given that \ilc{Gauge = empty} first, and again known that \ilc{Starts = no}; we may query on the probability of \ilc{Battery} and \ilc{Fuel}:
+
+\begin{lstlisting}
++---------------+----------------+          +-----------------+-------------+
+| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
++===============+================+          +=================+=============+
+| Battery(good) |         0.9410 |          | Fuel(not empty) |      0.0694 |
++---------------+----------------+          +-----------------+-------------+
+| Battery(bad)  |         0.0590 |          | Fuel(empty)     |      0.9306 |
++---------------+----------------+          +-----------------+-------------+
+\end{lstlisting}
+
+We may tell the leading potential cause of the problem might be \ilc{Fuel = empty}.
+Now we may ask our friend to collect information on \ilc{Turns Over} where we were given that \ilc{Turns Over = no}, we may again query on the probability of \ilc{Battery} and \ilc{Fuel}:
+
+\begin{lstlisting}
++---------------+----------------+          +-----------------+-------------+
+| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
++===============+================+          +=================+=============+
+| Battery(good) |         0.4726 |          | Fuel(not empty) |      0.5567 |
++---------------+----------------+          +-----------------+-------------+
+| Battery(bad)  |         0.5274 |          | Fuel(empty)     |      0.4433 |
++---------------+----------------+          +-----------------+-------------+
+\end{lstlisting}
+
+We observed that the probability of fuel being empty has dramatically decresed, and now the leading potential cause for the problem become \ilc{Battery = bad} -- and that's the true cause of the problem.
 

### 2021-03-09 A2/article.tex
✨ Add A2 Exploration on Boltzmann machine and its representation power.

@@
 
-\section*{Exploration}
+\section*{Exploration: Boltzmann machine, and it's representation power against sigmoid belief network.}
+
+I decided to take a slithly eccentric topic for this exploration as I don't really feel like designing a belief network, assigning some made up probablities, and calculate some conditional probablities.
+Since I feel like this will be very close to the pratice in \textbf{Question 5}, and with the help of \ilc{pgmpy}, a made up belief network by myself might be terrible to calculate by hand, but certainly not so hard to a computer; and it will be hard to show the effect of an optimization algorithm in a small scale belief network.
+
+However, I have spent quite amount of time in \textbf{Question 4} by exploring the representation power between the noisy-OR-based network and a sigmoid-based network; and the literatures I consulted dicussed some very interesting concepts regarding the \textit{Boltzmann Machine}, and its representation power in relationship to some more common belief networks -- like the sigmoid belief network.
+So I'd like to take this oppotunity to give a brief introduction of Boltzmann machine: what it is, how it work in the context of probability inference, and how it is applied to modern neural network learning.
+Then I will walk through mathematical examples analysing the representation power between a Boltzmann machine and a sigmoid belief network in the context of 2-, 3-, and 4-unit.
+Althought the networks presented here are not ``more complex'' in terms of number of variables comparing to other questions in this assignment, but it is certainly more involved as it is an analytical review of different learning model instead of calculating definitive answers; and for this I hope the grader won't consider this to be too shallow.\newline
+
+\paragraph{A quick introduction of Boltzmann machine.}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.4\linewidth]{{fig/bm}.png}
+    \caption*{Structure of Boltzmann machine}
+\end{figure}
+
+A Boltzmann machine is a type of stochastic recurrent neural network that is known to be widely applied to binary model inference problems.
+By inspecting the above diagram, it is clear that the Boltzmann machine is different to the neurol network (or deep belief network, to be precise) as it:
+
+\begin{itemize}
+    \item Has no output layer and it is undirected.
+    \item Has connections between input (in this case, the three blue) nodes.
+\end{itemize}
+
+The second property can be even more elaborate to the fact that every visible nodes within a Boltzmann machine can connect to each other, but this is merely a structure difference -- a restriced Boltzmann machine will not have this property, but still has no output nodes (and this will be a layer/block within deep neural network).
+The Boltzmann machine is an energy based system, where as each visible node will take a binary input, and by calculate the total energy of the system, we may infer on probablity of a particular visible node, or join distribution of couple visible nodes.
+
+The energy of the total system is denoted as $E$, where:
+
+\begin{equation*}
+    -E(v, h) = \sum_{i \in v} v_i b_i + \sum_{k \in h} h_k b_k + \sum_{i \leq j} v_i v_j w_{ij} + \sum_{i, k} v_i h_k w_{ik} + \sum_{k \leq l} h_k h_l w_{kl}
+\end{equation*}
+
+\begin{itemize}
+    \item $v$: visible nodes.
+    \item $h$: hidden nodes.
+    \item $b_i$: bias of node with index $i$, wheather this node is $v_i$ or $h_i$ is implied through mutiplication. i.e., $b_i$ in $h_i b_i$ is the bias of a hidden node $h_i$
+    \item $w_{ij}$: weight between nodes of index $i$ and $j$, wheather the nodes are visible or not (or one visible one hidden) is implied through mutiplication.
+\end{itemize}
+
+Since the energy formula is quite long, it might be hard to digest.
+But essentially, take an edge of the proposed network (say it is $e$) and mutiply it with two of its verticies (say it is $e_A, e_B$), keep the result and sum it up by doing this to every edges as $\sum\limits_{e \in E(G)} e e_A e_B$.
+Here's an example made by \href{https://www.youtube.com/watch?v=5jaBneYd5Ig}{Colin Reckons} (note we omitted connection between visible nodes for simplicity):
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.3\linewidth]{{fig/bm_example}.png}
+    \caption*{An example Boltzmann machine}
+\end{figure}
+
+Say we have $-E(<v_1, v_2, h_1, h_2>)$, the total energy of the network is:
+
+\begin{align*}
+    -E(<v_1, v_2, h_1, h_2>) &= w_{v_1 h_1}(v_1 \cdot h_1) - w_{h_1 h_2}(h_1 \cdot h_2) + w_{h_2 v_2}(h_2 \cdot v_2) =  2\\
+    -E(<1, 1, 1, 1>) &= 2(1 \cdot 1) - 1(1 \cdot 1) + 1(1 \cdot 1) =  2\\
+    -E(<1, 1, 1, 0>) &= 2(1 \cdot 1) + 1(1 \cdot 0) + 1(0 \cdot 1) = 2 \\
+    &\dots
+\end{align*}
+
+We denote this total energy as $-E(v, h)$ to represent it is the total energy among all visible and hidden nodes; and we may infer the probablity of $p(v, h)$ and $p(v)$ as:
+
+\begin{align*}
+    p(v, h) &= \frac{e^{-E(v, h)}}{\sum_{u, g} e^{-E(u, g)}} \\
+    p(v) &= \frac{\sum_h e^{-E(v, h)}}{\sum_{u, g} e^{-E(u, g)}}
+\end{align*}
+
+Where $u, g$ represents all possible combination of vissible and hidden nodes.
+In this example, we have $<v_1, h_2, h_1, h_2>$ to be a 4-bit binary vector, so there can be 16 possible combinations.
+Thus, we may proceed and infer the possible of all conditions as:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{{fig/bm_result}.png}
+    \caption*{Probablity inferences of the example Boltzmann machine}
+\end{figure}
+
+Note this \textcolor{red}{39.70} is simply the sum of all $e^{-E}$, and each $p(v, h)$ are the $e^{-E}$ on its row divided by this 39.70.
+And the four $p(v)$ are sum of all $p(v, h)$ with the same $v$ inputs.\newline
+
+\paragraph{Representation power on 2- and 3-unit binary input.}
+Now we want to show that \textbf{ANY} distribution of a two-unit binary input (with excaption of having extreme output like $0, 1$, which requires a sum-of-weighted-inputs of $\pm infty$) can be represented by the Boltzmann machine and also a sigmoid feedforward network.
+We start with the sigmoid feedforward network.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{{fig/2_unit}.png}
+    \caption*{A Sigmoid feedforward network that can represent any distribution of a two-unit binary input.}
+\end{figure}
+
+It is easy to tell that this network can represent any distribution of a two-unit binary input, as we have $P(Y \mid X=1)$ and all possible $P(X)$s, all we have to do is to make it proportional to the sigmoid transformation so that $\sigma(z)$ can be the desired probablity output (where $z$ is the sum of weighted inputs).
+\newline
+
+Now to represent the same distribution with a Boltzmann machine.
+We have the energy of $<0, 0>$ to be always zero, so the $p(<0, 0>)$ must relay on the its proportion to the sum of all energy.
+For $<1, 0>$ and $<0, 1>$, we may simply set the desired probablity distribution to be proportional to bias of each $1$ node, as the total energy in such state is always $e^{-B_x}$ (assuming $x$ is the node with $1$).
+Last, for the state $<1, 1>$, we may set the desired probablity to be proportional to the weight $w_{xy}$.
+As this Boltzmann machine can represent the distributions all four states of a 2 unit binary input with the same bias and weights, it can also represent any distribution of a two-unit binary input just like the sigmoid feedforward network.
+
+This inference can be further elaborated to three-unit binary input as showned in \href{http://www.cs.toronto.edu/~bonner/courses/2016s/csc321/readings/Connectionist%20learning%20of%20belief%20networks.pdf}{R.M.
+Neal, Connectionist learning of belief networks, Artificial Intelligence 56 (1992) 71–113.} by doing what we did first for the first to node, and connect to third node with weights of its conditional probablity to the first two nodes.
+However, it can only be shown a distribution that is representable in sigmoid feedforward network is equally representable in a Boltzmann machine and vice versa.
+This is not to say that any \textbf{ANY} (non-extreme, as for no $0, 1$ probablity outputs) distribution of a three-unit binary input can be represent in a sigmoid feedforward network / Boltzmann machine, as in \textbf{Question 4.4.} we have shown a three-unit system with noisy-OR distribution can't be represented by a sigmoid feedforward network.
+
+\paragraph{Representation power on 4-unit binary input.}
+
+So far, although we have demonstrated that there are (non-extreme) distribution of 3-unit binary input that cannot be represented by a sigmoid network or a Boltzmann machine, the two still got the same representation power.
+But this is not the case when we raise to 4-unit binay input, as each of them got some distribution the other can't represent.
+Due to time and space constraint, here's a Boltzmann machine that can't be represent by a sigmoid netwrok, designed by R.M.
+Neal:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.3\linewidth]{{fig/bm_no_sig}.png}
+    \caption*{A Boltzmann machine that is untranslatable to sigmoid belief network}
+\end{figure}
+
+Since the Boltzmann machine is symmetric, we mays start with any node and call this unit 4.
+For the weights that going into the second-last unit (unit 3 ), due to the symmetry shape, the two weights must be equal (we denote them as $w$, and let the bias of unit 3 to be $b$).
+Now in the case of unit 3 and unit 4 have value of 1s, and unit 1 and unit 2 have zero or one or two ones among them, we must have the following relationship (LHS Boltzmann and RHS sigmoid):
+
+\begin{align*}
+    e^{1} &= \propto e^b \frac{\sigma(1)}{\sigma{0}} &\text{zero 1 in unit 1\&2} \\
+    e^{2} &= \propto e^{b + w} \frac{\sigma(2)}{\sigma{1}} &\text{one 1 in unit 1\&2} \\
+    e^{3} &= \propto e^{b + 2w} \frac{\sigma(3)}{\sigma{2}} &\text{two 1s in unit 1\&2}
+\end{align*}
+
+Which is clear that this is not a solvable linear system, and therefore this distribution cannot be translate to a sigmoid belief network.
 

### 2021-03-30 A3/code/e1.py
⚙️ Add code and source files for A3.

@@ -1,0 +1,119 @@
+import cv2
+import numpy as np
+import matplotlib.image as cv
+import matplotlib.pyplot as plt
+import copy
+
+def bw_image(path = 'A3/code/hello_world.png'):
+    img = cv.imread(path)
+    img = img.copy()
+    img[img < 0.5] = -1
+    img[img >= 0.5] = 1
+    img = img[: , :, 0]
+    return img
+
+
+def add_noise(img, thlr = 0.95):
+    m, n = np.shape(img)
+    noised = np.copy(img)
+    for i in range(m):
+        for j in range(n):
+            if np.random.rand() >= thlr:
+                noised[i][j] = -noised[i][j]
+    return noised
+
+
+
+def neighbor_sum(x, i, j, more_flag = False):
+    def get_pix(x, i, j):
+        try:
+            return x[i][j]
+        except IndexError:
+            return 0
+    if not more_flag:
+        sum = x[i][j] * (get_pix(x, i+1, j) + get_pix(x, i-1, j) + get_pix(x, i, j+1) + get_pix(x, i, j-1))
+    else:
+        sum = x[i][j] * (get_pix(x, i+1, j) + get_pix(x, i-1, j) + get_pix(x, i, j+1) + get_pix(x, i, j-1)
+                    + get_pix(x, i+1, j+1) + get_pix(x, i-1, j-1) + get_pix(x, i-1, j+1) + get_pix(x, i+1, j-1)
+                    + get_pix(x, i+2, j) + get_pix(x, i-2, j) + get_pix(x, i, j+2) + get_pix(x, i, j-2))
+    return sum
+
+
+def denoise(img, h=5, beta=5, eta=5, energy_plot = False, energy_cap = False, neighbor_eval = False):
+    m, n = np.shape(img)
+    x = np.copy(img)
+    y = np.copy(img)
+
+    pass_counter = 0
+    energy_change = []
+    current_energy = 0
+    if energy_plot:
+        current_energy = init_energy(img, h, beta, eta)
+        energy_change.append(current_energy)
+
+
+    for i in range(m):
+        for j in range(n):
+            x_k = x[i][j]
+            first_term = -2 * h * x_k
+            second_term = 4 * beta * neighbor_sum(x, i, j)
+            third_term = 2 * eta * x_k * y[i][j]
+            neighbor_term = 4 * beta * neighbor_sum(x, i, j, more_flag = True)
+            if not neighbor_eval:
+                delta = first_term + second_term + third_term
+            else:
[…]

### 2021-03-30 A3/article.tex
✨ Add answers and exploration for A3.

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\newcommand{\p}{\partial}
\usepackage{etoolbox}
\usepackage{algorithm}
\usepackage{changepage}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}

\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
% \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\providecommand{\myceil}[1]{\left \lceil #1 \right \rceil }
\providecommand{\myfloor}[1]{\left \lfloor #1 \right \rfloor }

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\begin{document}

\title{\textbf{CSDS 491: Assignment 3}}

\author{Shaochen (Henry) ZHONG, \ilc{sxz517@case.edu}}

\date{Due and submitted on 03/29/2021 \\ Spring 2021, Dr. Lewicki}
\maketitle

% We've received some common questions for A1.  Here are some hints to help you go in the right direction.
%
% Q2.2: You need only prove there is a logical gap that would require an additional assumption.
%
% Q4: It is helpful to use the conjugate prior as shown in lecture.  The  normalization factor for the posterior then has the same form as the prior (but with different arguments).
%
% Q5: Remember that the data consists of event times, and the posterior is in terms of the Poisson rate parameter lambda given the number of observed data events in a given duration from 0 to T.  You only need a single Poisson likelihood combined with the prior to get the posterior.
%
% Exploration: The terms discrete and continuous refer to the variables in the model.  See the rubric for the grading criteria. Try to make your inference problems simple models of real-world situations, rather than just numerical examples.

\section*{E1. MRFs and Images Denoising}

\subsection*{1.1.}

Per question, known that the total energy of the system is:

\begin{equation*}
    E(x, y) = h \sum_{i} x_i - \beta \sum_i \sum_{j \in ne(x_i)} x_i x_j - \eta \sum_i x_i y_i
\end{equation*}

By ``letting out'' $x_k$, we have the following:

\begin{align*}
    E(x, y) &= (h \sum_{i / k} x_i - \beta \sum_{i / k} \sum_{j \in ne(x_i) / k} x_i x_j - \eta \sum_{i / k} x_i y_i) + h \cdot x_k - \beta \sum_{j \in ne(x_k) }x_k x_j - \beta \sum_{j \in ne(x_k) } x_j x_k- \eta x_k y_k \\
    &= C + h \cdot x_k - \beta \sum_{j \in ne(x_k) }x_k x_j - \beta \sum_{j \in ne(x_k)}  x_j x_k- \eta x_k y_k
\end{align*}

On the other hand, assume $x$ is now $\overline{x}$ with $x_k$ flipped, we must have:

\begin{align*}
    E(\overline{x}, y) &= (h \sum_{i / k} x_i - \beta \sum_{i / k} \sum_{j \in ne(x_i) / k} x_i x_j - \eta \sum_{i / k} x_i y_i) + h \cdot x_k - \beta \sum_{j \in ne(x_k) }x_k x_j - \beta \sum_{j \in ne(x_k) } x_j x_k- \eta x_k y_k \\
    &= C + h \cdot (-x_k) - \beta \sum_{j \in ne(x_k)} (-x_k) x_j - \beta \sum_{j \in ne(x_k) }x_j (-x_k) - \eta (-x_k) y_k
\end{align*}

Thus, the change in energy will be:

\begin{align*}
    \Delta E &= E(\overline{x}, y) - E(x, y) \\
    &= C - C -2 h \cdot x_k -4\beta \sum_{j \in ne(x_k)} (-x_k) x_j + 2 \eta x_k y_k \\
    &= -2 h \cdot x_k + 4\beta \sum_{j \in ne(x_k)} x_k x_j + 2 \eta x_k y_k
\end{align*}

\subsection*{1.2.}

You may refer to \ilc{A3/code/e1.py} for my code, the method for energy updating calculation is \ilc{denoise()}. Currently my method is ``updated'' for tasks of below questions so it might be a little complicate to read, but the generally idea is to:

\begin{itemize}
    \item Pick one pixel
    \item Calculate the energy change $\Delta E$ in terms of first (with parameter $h$), second (with parameter $\beta$), and third (with parameter $\eta$) terms per each pixel.
    \item Check if the energy change $\Delta E$ is negative, if so, flip the pixel. This is because lower energy yields a higher probablity $p(x, y)$, so the pixel-just-flipped is more likely to be a noise pixel.
    \item Proceed to the next pixel. In my case I just loop through every pixel along their indicies.
\end{itemize}

\subsection*{1.3.}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\linewidth]{{fig/E1.3}.png}
\end{figure}
As whenever the $\Delta E$ is negative, we take the flip, the overall energy is decreasing through out the whole process as the passing number increasing.

\subsection*{1.4.}

\begin{figure}[H]
\minipage{0.25\textwidth}
  \includegraphics[width=\linewidth]{{fig/E1.4}.png}
  \caption*{Original Image}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.4.1}.png}
    \caption*{Noised Image}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.4.2}.png}
    \caption*{50\% Denoised}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.4.3}.png}
    \caption*{Fully Denoised}
\endminipage
\end{figure}

Note since I iterate through all pixels along their indicies, so the 50\% image seems only ``top half denoised.'' If I use random order evaluate all pixels, there won't be such distinction.

\subsection*{1.5.}

So we have three paramaters to evaluate. The $h$ is regarding the original value of pixel-in-question, the $\beta$ is regarding the neighbors of the pixel-in-question, and the $\eta$ is about the change of pixel-in-question (if flipped, energy decrease, otherwise increase).

It is my understanding that $h$ serves as a bias of the model, where as it prefer a pixel value more than the other. In the case that $h > 0$, the model will prefer pixels with $1$-value (white); and if $h < 0$, the model will prefer pixel with $-1$-value (black). We may confirm this theory by comparing the following trials:

\begin{figure}[H]
\minipage{0.25\textwidth}
  \includegraphics[width=\linewidth]{{fig/E1.4}.png}
  \caption*{Original Image}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.5.1}.png}
    \caption*{$h = -2, \beta = 2, \eta = 2$}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.5.2}.png}
    \caption*{$h = 2, \beta = 2, \eta = 2$}
\endminipage
\end{figure}

where the $h>0$ result contains more white pixels to $h<0$ trial.\newline

\noindent The $\beta$ term is about the neighbors of a pixel. With $\beta$ being larger, the resultant picture will look more ``clustered'' pixel nearby will tend to be the same color.

\begin{figure}[H]
\minipage{0.25\textwidth}
  \includegraphics[width=\linewidth]{{fig/E1.5.3}.png}
  \caption*{$h = 0, \beta = -2, \eta = 2$}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.5.4}.png}
    \caption*{$h = 0, \beta = 0, \eta = 2$}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.5.5}.png}
    \caption*{$h = 0, \beta = 2, \eta = 2$}
\endminipage
\end{figure}

It is clear that with $\beta < 0$, the model tend to believe there's no consistancy between nearby pixels and therefore more ``noisy''. With $\beta = 0$ and no $h$ bias, since there is no flip of pixel (and thus no $\eta$ term decrease), the model will not conduct any denoise operation. With $\beta > 0$, we have a meaning denoised result as expected.\newline

\noindent Last, the $\eta$ term represent the relationship between the noised image and the original image. As we perviously mentioned, such term will decrese if a flip occurs and increase if the value is kept; so a large $\eta$ will encourage the model to kept the pixel value of the noised image. This is probably only good if the original image has minimum amount of noise added.

\begin{figure}[H]
\minipage{0.25\textwidth}
  \includegraphics[width=\linewidth]{{fig/E1.5.6}.png}
  \caption*{$h = 0, \beta = 2, \eta = -2$}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.5.7}.png}
    \caption*{$h = 0, \beta = 2, \eta = 0$}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.5.8}.png}
    \caption*{$h = 0, \beta = 2, \eta = 10$}
\endminipage
\end{figure}

We may observe that with $\eta < 0$, flipping is encouraged so a lot of information became missing. As $\eta$ increase, the model became reluctant to flip and therefore many noisy pixels remain.

\subsection*{1.6.}

I'd like my new energy function to not only consider the four direct neighbors of the pixel-in-question, but also its four other ``corner neighbors.'' On top of this, I'd like each pixel-in-question to expand one extra ``layer'' of scoupe when evaluting neighbors, as it will take the more distant neighbor of neighbors into consideration.

\begin{align*}
    E(x, y) = h \sum_{i} x_i - \beta \sum_i \sum_{j \in around(x_i)} x_i x_j - \beta \sum_i \sum_{k \in outer(i)} x_i x_k - \eta \sum_i x_i y_i \\
    \Delta E = -2 h \cdot x_k + 4\beta \sum_{j \in around(x_k)} x_k x_j + 4\beta  \sum_i \sum_{k \in outer(i)} x_i x_k + 2 \eta x_k y_k
\end{align*}

Since the original energy function only consider four of the neighbors, it has the potential of offering less desired performances with images of interlaced patterns. This is because the model will only exam the four closest neighbors of each pixel; with images with interlaced patterns it might have trouble of distinguishing actual information verses a large grain noise.

\begin{figure}[H]
\minipage{0.25\textwidth}
  \includegraphics[width=\linewidth]{{fig/E1.6}.png}
  \caption*{Input}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.6_noised}.png}
    \caption*{Noised}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.6.1}.png}
    \caption*{Original $\Delta E$}
\endminipage\hfill
\minipage{0.25\textwidth}
    \includegraphics[width=\linewidth]{{fig/E1.6.2}.png}
    \caption*{Updated $\Delta E$}
\endminipage
\end{figure}

This might be a set that is harder to observe the difference. But if you zoom in you will realize the updated $\Delta E$ is more ``connected'' at the end of each edge, which is a better presentation of the original input. This is due to the updated $\Delta E$ takes more and farther nearby pixels into consideration.

\section*{E2. Graphical Representation}

\subsection*{2.1.}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.3\linewidth]{{fig/E2.1}.png}
\end{figure}

For Bayes net, we have:

\begin{equation*}
    p(a, b, c, d, e, f, g) = p(a)p(b)p(c \mid a)p(d \mid a, b)p(e \mid b)p(f \mid c)p(g \mid d)p(h \mid d)
\end{equation*}

\noindent For MRF, we connect the parents and have:
\begin{equation*}
    p(a, b, c, d, e, f, g) = \psi(a, c)\psi(a, b, d)\psi(b, e)\psi(c, f)\psi(d, g)\psi(d, h)
\end{equation*}
\subsection*{2.2.}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.3\linewidth]{{fig/E2.2}.png}
\end{figure}

THe joint distribution will be:

\begin{equation*}
    p(a, b, c, d, e, f, g) = f_1(a, c)f_2(a, b, d)f_3(b, e)f_4(c, f)f_5(d, g, h)
\end{equation*}

\subsection*{2.3.}

\begin{figure}[H]
\minipage{0.3\textwidth}
  \includegraphics[width=\linewidth]{{fig/E2.3.1}.png}
  \caption*{Graph 1}
\endminipage\hfill
\minipage{0.3\textwidth}
    \includegraphics[width=\linewidth]{{fig/E2.3.2}.png}
    \caption*{Graph 2}
\endminipage
\end{figure}

For \textbf{Graph 1}, we added factor functions between all connections, and the probablity function will be:

\begin{align*}
    f_1(S, R, T) &= p(T \mid S, R)p(S)p(R) \\
    f_2(R, J) &= p(J \mid R) \\
    p(S, R, T, J) &= f_1(S, R, T)f_2(R, J) = p(T \mid S, R)p(S)p(R)p(J \mid R)
\end{align*}

On ther other hand for \textbf{Graph 2} we added factor functions for every nodes, thus:

\begin{align*}
    f_1(S) &= p(S) \\
    f_2(R) &= p(R) \\
    f_3(S, R, T) &= p(T \mid S, R) \\
    f_4(R, J) &= p(J \mid R) \\
    p(S, R, T, J) &= f_1(S)f_2(R)f_3(S, R, T)f_4(R, J) = p(S)p(R)p(T \mid S, R)p(J \mid R)
\end{align*}

\section*{E3. The Sum Product Algorithm}

\subsection*{3.1.}

\begin{align*}
    \mu_{a \to f_1} &= \mu_{b \to f_1} = \mu_{d \to f_2} = 1 \\
    \mu_{f_1 \to c} &= \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \mu_{b \to f_1} \\
    \mu_{f_2 \to c} &= \sum_d f_2(c, d)\mu_{d \to f_2}
\end{align*}

\subsection*{3.2.}

\begin{align*}
    p(c) &= \mu_{f_1 \to c}\mu_{f_2 \to c} \\
    &=  \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \mu_{b \to f_1} \cdot \sum_d f_2(c, d)\mu_{d \to f_2} \\
    &= \sum_a \sum_b f_1(a, b, c)  \sum_d f_2(c, d)
\end{align*}

\subsection*{3.3.}

\begin{align*}
    p(c) &= \sum_a \sum_b \sum_d p(a, b, c, d) \\
    &= \sum_a \sum_b \sum_d f_1(a, b, c)  \sum_d f_2(c, d) \\
    &= \sum_a \sum_b f_1(a, b, c)  \sum_d f_2(c, d)
\end{align*}

\subsection*{3.4.}

% \begin{align*}
%     p(c) &= \Pi_{s \in ne(c)} \mu_{f_s \to c} \\
%     &= \mu_{f_1 \to c}\mu_{f_2 \to c} \\
%     &= \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \mu_{b \to f_1} \cdot \sum_d f_2(c, d)\mu_{d \to f_2} \\
%     &= \sum_a \sum_b f_1(a, b, c)  \sum_d f_2(c, d)
% \end{align*}

The information passing of graph-in-question is defined as:

\begin{align*}
    \mu_{a \to f_1} &= \mu_{d \to f_2} = 1 \\
    \mu_{b \to f_1} &= \mu_{f_2 \to b} \\
    \mu_{b \to f_2} &= \mu_{f_1 \to b} \\
    \mu_{f_1 \to b} &= \sum_a \sum_c f_1(a, b, c) \mu_{a \to f_1} \mu_{c \to f_1} \\
    \mu_{f_1 \to c} &= \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \mu_{b \to f_1} \\
    \mu_{f_2 \to b} &= \sum_c \sum_d f_2(b, c, d) \mu_{c \to f_2} \mu_{d \to f_2} \\
    \mu_{f_2 \to c} &= \sum_b \sum_d f_2(b, c, d) \mu_{b \to f_2} \mu_{d \to f_2}
\end{align*}

Now we have:

\begin{align*}
    p(c) &= \mu_{f_1 \to c} \mu_{f_2 \to c} \\
    &= \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \mu_{b \to f_1} \cdot \sum_b \sum_d f_2(b, c, d) \mu_{b \to f_2} \mu_{d \to f_2} \\
    &= \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \mu_{f_2 \to b} \cdot \sum_b \sum_d f_2(b, c, d) \mu_{d \to f_2} \mu_{f_1 \to b}   \\
    &=  \sum_a \sum_b f_1(a, b, c) \mu_{a \to f_1} \sum_c \sum_d f_2(b, c, d) \mu_{c \to f_2} \mu_{d \to f_2} \cdot \sum_b \sum_d f_2(b, c, d) \mu_{d \to f_2} \sum_a \sum_c f_1(a, b, c) \mu_{a \to f_1} \mu_{c \to f_1} \\
    &=  \sum_a \sum_b f_1(a, b, c) \sum_c \sum_d f_2(b, c, d) \mu_{c \to f_2} \cdot \sum_b \sum_d f_2(b, c, d)  \sum_a \sum_c f_1(a, b, c) \mu_{c \to f_1}\\
    &= \sum_a \sum_b f_1(a, b, c)  \mu_{f_2 \to b} \cdot \sum_b \sum_d f_2(b, c, d) \mu_{f_1 \to b} \\
    &...
\end{align*}

The equation will eventually include itself and therefore become an infinite loop.

\section*{Exploration}

For this exploration I'd like to explore more about image denoising. I will first talk about my experience with \textit{image stacking}, a popular image denoising method used by modern photographer. Then I will discuss paper \href{https://arxiv.org/abs/1901.06046}{Good Similar Patches for Image Denoising} by Lu.\newline

\noindent My dad and I are both photographers, and I was lucky enough to have an early and healthy exposure to many photography gears and concepts. Ever since I was semi-fluent on computers, Photoshop become one of my favourite thing to study on; and I have spent much of my middle/highschool life reading random tutorials online to learn a new ``trick'' or two. And one day I came across the concept of \textit{image stacking}, although I haven't realized much back then, that might be my first ``relatively in-depth'' experience to image denoising.

Modern landscape photography tends to push cameras into their limits, mostly because of the popularity of taking images with high dynamic range -- which can be roughly understood as the brightness difference between the lightest and the darkest pixel within the same image. Because image with high dynamic range are friendly to post-processing and the retouched version of such images may have the ``HDR-like'' affect that is very catchy to audiances (we typically can't capture such large range of brightness at the same time with raw eyes).

\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{{fig/E3.1}.jpg}
    \caption*{The Dunhuang frescoes, by Thomas Chu}
\end{figure}

However, high dynamic range photography has one intrinsic problem to solve: the darker side of the image are often ``under exposure'' and therefore prone to noises and information losing. And there are three typical solutions to this problem: ND filters, HDR, or image stacking. The first two are essentially the same solution, as we simply ``lighten up'' the darker side of the image to gain more information.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\linewidth]{{fig/E3.2}.jpg}
    \caption*{ND Filter demo}
\end{figure}

With the difference being ND filter does it with one-take, but HDR requires multiple exposures (with different light flux setting, usually controlled by shutter speed) and combining multiple images into one in post processing. We may say the apporach of these two methods is to proactively eliminate noise to find the \textit{ground truth} version of photo. But sometime these two methods are infesible, e.g., the brightness difference (\textit{stops}) of your ND filter is not high or sutiable enough for the scene, there are moving object in the scene so you can't easily combine two images together or the noise is still high after such approach, etc. And image stacking may come at your rescue.\newline

The difficulty of real word image denoising task is there is no \textit{ground truth} of a photo. As every photo you took contain a certain amount of noise, and there's no way you can idealy quantify the noisy-level of an image. So the philosophy of image stacking is simply do ``repetition,'' where you record the same scene with the same setting for a large amount of times. As (a portion of) noise are essentially randomly captured, with repetition, we may identify the noisy-pixels and patch them with pixels from other images. This brings superior performance, as you can always push for higher SNR with more repetition and you got mutiple chance to identify a noise and select patches. It is also friendly to moving objects, as you got a continuous series of images to ``blur it out'' and pratically, it is easy to execuate by leaving your camera on a tripod and wait.

\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{{fig/E3.3}.jpg}
    \caption*{Image Stacking demo, visible improvement on the RHS with more images being stacked}
\end{figure}

\noindent So image stacking is an effective method because it can easily identify the noisy-pixel, which will be the most deviated pixel among the stack (on the same coordiate); and it is straighforward on finding patch, as it will be the average of pixel which we deemed to be not noise. In \textbf{Exercise 1} of this assignment, we only focused on identifying noise due to the nature of the input image is binary. But in real word application, noise are often relatively easy to identify -- as you can easily recognize a chroma noise out of an image -- but what to patch the noise with, is often regarded as a harder question to answer.

In this assignment, we center our patch searching as the noisy-pixel, where we evaluate the neighbors of the noisy-pixel to determine the patch -- this is regarded as \textit{Nearest Neighbour Search (NNS)}. However, in non-binary setting, the neighbors of a noisy-pixel may (and often are) under influence of other noise, and therefore renders a patch which is far from the ground truth of the image.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{{fig/E3.4}.png}
    \caption*{In this case, the NNS candidate (blank square) to the noisy black circle is far away from the ground truth patch (black square)}
\end{figure}

To address this problem, Lu proposed a method to select patch by conduct clustering on candidate patch groups.

\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{{fig/E3.5}.png}
    \caption*{Lu's method of patch selection}
\end{figure}

The takeaway of Lu's approach is to not to center the path search among the noisy reference, but to cluster all nearby pixels into different patch groups first. Then measure the distance between the noisy reference to center of different patch groups, and select patch from the patch group with closest distance.

Lu introduced that each patch group can be modeled as a Gaussian function, and the whole set of candidate patches can be approximated as a Gaussian Mixture Model. The clustering will follow the the Minimum Decription Length criteria:

\begin{equation*}
    MDL (K, \theta) = - \log p_{Q}(Q \mid K, \theta) + \lambda L \log(mn)
\end{equation*}

The first term is defined as:

\begin{equation*}
    - \log p_{Q}(Q \mid K, \theta) = \sum_{k = 1}^K \sum_{j = 1}^{m_k} || q_j - u_k ||^2
\end{equation*}

% L &= K(1 + n + \frac{(n +1)n}{2}) - 1
where $m_k$ and $\mu_k$ are the number of patches in the $k$-th cluster and its centroid. Since the goal is to minimize the $MDL (K, \theta)$, clusters with small ``diameter'' and therefore smaller intra-patch distances are prefered. Note here $Q$ is a set of $m$ \ $n-$dimensional candidate patches, and $\theta$ is the paramaters of GMM.

\begin{equation*}
    L = K(1 + n + \frac{(n +1)n}{2}) - 1
\end{equation*}

The author refered the second term of his $MDL (K, \theta)$ equation as a regularization term to punish large clusters, and $\lambda$ to be a paramater that balances the contribution of two terms. Well I don't quite understand why would he make such complicate of a $L$ since $m$ directly defines the size of a cluster; it seems $L$ is doubled as a definition of number of parameters used in $\theta$ (The author also shows with optimization of K-means++ on clustering, the regularization term will be a constant and can be ignored). After all the goal is to have many small clusters (in terms of diameters and sizes), so that we may have more candidate to attribute our noisy reference to. And intuitively, such clustering will be more fine grained and therefore result in better patch selection: the relationship the ground trutch patch and the selected patch will be closer (as they are in a small cluster), and therefore a better patch can be selected if we may attribute the noisy reference to the right cluster. .\newline

The author then go on and optimized his algorithm with \textit{unrelisable pixel detection} algorithm to detect the noisy reference. He further testified his method -- as a plug of many classic patch-based denoising algorithms -- and demonstrated state-of-the-art performance on some typical image denoising dataset. Since this is my only exposure to image denoising paper, I can't offer much insight about its performance in comparision to the others. Judging from the demo images in paper, the improvement does seem significant.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\linewidth]{{fig/E3.6}.png}
    \caption*{\textit{BST} shows improvement on three different denoising algorithms}
\end{figure}

\end{document}

### 2021-04-19 A4/code/e2.py
⚙️ Add code for A3 E2.

@@ -1,0 +1,44 @@
+# Ref. https://stackoverflow.com/questions/20011122/fitting-a-normal-distribution-to-1d-data
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.stats import norm
+
+
+N = 999999
+
+
+X = norm.rvs(0, 3, size = N)
+Z = norm.rvs(1, 4, size = N)
+Y = norm.rvs(1, 5, size = N)
+
+X_Z = X + Z
+
+plt.subplot(1,2,1)
+mu, std = norm.fit(Y)
+plt.hist(Y, bins=25, density=True, alpha=0.6, color='b')
+xmin, xmax = plt.xlim()
+x = np.linspace(xmin, xmax, 100)
+p = norm.pdf(x, mu, std)
+plt.plot(x, p, 'k', linewidth=2)
+title = "Fit results: mu = %.2f,  std = %.2f" % (mu, std)
+plt.title(title)
+plt.ylabel('p(y)')
+plt.xlabel('y')
+
+
+plt.subplot(1,2,2)
+mu, std = norm.fit(X_Z)
+plt.hist(X_Z, bins=25, density=True, alpha=0.6, color='g')
+xmin, xmax = plt.xlim()
+x = np.linspace(xmin, xmax, 100)
+p = norm.pdf(x, mu, std)
+plt.plot(x, p, 'k', linewidth=2)
+title = "Fit results: mu = %.2f,  std = %.2f" % (mu, std)
+plt.title(title)
+plt.ylabel('p(x + z)')
+plt.xlabel('x + z')
+
+
+plt.show()

### 2021-04-19 A4/article.tex
✨ Add answer to A3 E2.

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\newcommand{\p}{\partial}
\usepackage{etoolbox}
\usepackage{algorithm}
\usepackage{changepage}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}

\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
% \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\providecommand{\myceil}[1]{\left \lceil #1 \right \rceil }
\providecommand{\myfloor}[1]{\left \lfloor #1 \right \rfloor }
\providecommand{\qbm}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\providecommand{\qpm}[1]{\begin{pmatrix} #1 \end{pmatrix}}
\providecommand{\norm}[1]{\left\lVert #1 \right\rVert}
\providecommand{\len}[1]{\left| #1 \right|}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\begin{document}

\title{\textbf{CSDS 491: Assignment 4}}

\author{Shaochen (Henry) ZHONG, \ilc{sxz517@case.edu}}

\date{Due and submitted on 04/21/2021 \\ Spring 2021, Dr. Lewicki}
\maketitle

\section*{E1. Multivariate Gaussians}
\section*{E2. Linear Gaussian Models}

\subsection*{2.1.}
% https://math.stackexchange.com/questions/60911/multivariate-normal-difference-distribution
% https://stats.stackexchange.com/questions/9879/addition-of-multivariate-gaussians

Known the characteristic functions of $x$ and $z$ is:

\begin{align*}
    \phi_x(u) &= \exp(i u^T \mu_x - \frac{1}{2} u^T \Sigma_x u) \\
    \phi_z(u) &= \exp(i u^T \mu_z - \frac{1}{2} u^T \Sigma_z u)
\end{align*}

Since $x$ and $z$ are independent, their sum will still conform to the normal distribution. We may confirm it by analysing the characteristic functions of $x + z$, which is $\phi_{x + z}(u) = E(e^{it(x + z)}) = E(e^{itx}) \cdot E(e^{itz}) = \phi_{x}(u) \phi_{z}(u)$, then we have:

\begin{align*}
    \phi_{x + z}(u) &= \phi_{x}(u) \phi_{z}(u) = \exp(i u^T \mu_x - \frac{1}{2} u^T \Sigma_x u) \cdot \exp(i u^T \mu_z - \frac{1}{2} u^T \Sigma_z u) \\
    &= \exp(i u^T (\mu_x + \mu_z) - \frac{1}{2} u^T (\Sigma_x + \Sigma_z) u) \\
    \Longrightarrow \phi_y(u) &= \exp(i u^T \mu_y - \frac{1}{2} u^T \Sigma_y u)
\end{align*}

Thus, we have $p(y) = \mathcal N(\mu_x+\mu_z,\Sigma_x+\Sigma_z)$.

\subsection*{2.2.}

% https://stats.stackexchange.com/questions/30588/deriving-the-conditional-distributions-of-a-multivariate-normal-distribution
% http://www.math.chalmers.se/~rootzen/highdimensional/SSP4SE-appA.pdf
% http://users.stat.umn.edu/~helwig/notes/norm-Notes.pdf
% https://www.coursera.org/lecture/linear-models-2/normal-conditional-distributions-LiHmE

For clarity, we let $p(x, y) = \mathcal N(\mu, \Sigma)$, where:

\begin{align*}
    \mu &= \qbm{\mu_x \\ \mu_y} \\
    \Sigma &= \qbm{\Sigma_{xx} & \Sigma_{xy} \\ \Sigma_{yx} & \Sigma_{yy}}
\end{align*}

Note each of $\Sigma{ab}$ is a matrix of $a \times b$.

We further know that there must be $p(y \mid x) = \mathcal N(\overline{\mu}, \overline{\mu})$, where:

\begin{align*}
    \overline{\mu} &= \mu_y + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) \\
    \overline{\Sigma} &= \Sigma_{yy} - \Sigma_{yx} \Sigma_{xx}^{-1}\sigma{yx}
\end{align*}

I don't know if we are expected to show the proof, but here's a trick that can do it rather quickly. Assume $z = y + Ax$ (not to be confused with the $z$ in \textbf{Question 2.1}) with $A = -\Sigma_{yx} \Sigma_{xx}^{-1}$, we have:

\begin{align*}
    cov(z, x) &= cov(y + Ax, x) = cov(y, x) + A \cov(x, x) \\
    &= \Sigma_{yx} + (-\Sigma_{yx} \Sigma_{xx}^-1) \Sigma_{xx} \\
    &\text{Now to deduce $E(y \mid x)$} \\
    \Rightarrow &\begin{cases}
        E(z | x) &= E(z) = E(y + Ax) = \mu_y + A \mu_x = \mu_y -\Sigma_{yx} \Sigma_{xx}^{-1} \mu_x \\
        E(z | x) &= E(y + Ax \mid x) = E(y \mid x) + A E(x \mid x)
    \end{cases} \\
    \Longrightarrow E(y \mid x) &= E(z | x) - A E(x \mid x) = \mu_y -\Sigma_{yx} \Sigma_{xx}^{-1} \mu_x - (-\Sigma_{yx} \Sigma_{xx}^{-1}) x \\
    &= \mu_y + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) = \overline{\mu}
\end{align*}

Similarily, we may find out $\overline{\Sigma}$ as (note that $cov(x, y) = cov(y, x)$ for scalar):

\begin{align*}
    var(y \mid x) &= var(z) = var(y + Ax) \\
    &= var(y) + A var(x) A' + A cov(y, x) + cov(x, y) A' \\
    &= \Sigma_{yy} + \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xx} \Sigma_{xx}^{-1} \Sigma_{xy} - \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{yx} - \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy} \\
    &= \Sigma_{yy} + \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy} - 2 \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy} \\
    \Longrightarrow \overline{\Sigma}&= \Sigma_{yy} - \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy}
\end{align*}

As know from \textbf{Exercise 2.1} that $p(y) = \mathcal N(\mu_x+\mu_z,\Sigma_x+\Sigma_z)$ and $p(x) = \mathcal N(\mu_x, \Sigma_x)$, we have $p(y \mid x) = \mathcal N(\overline{\mu}, \overline{\mu})$, where:

\begin{align*}
    \overline{\mu} &= \mu_y + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) = \mu_x \mu_z + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) \\
    \overline{\Sigma} &=  \Sigma_{yy} - \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy} = \Sigma_{x} + \Sigma_{z} -\Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy}
\end{align*}

\subsection*{2.3.}

According to \textbf{Exercise 2.1}, for $y = x + z$, we have $y \sim \mathcal N(\mu_x+\mu_z,\Sigma_x+\Sigma_z)$. For the ease of demo, let $\sigma_x = 3$ and $\sigma_z = 4$ so that we may have the nice $\sigma_x^2 + \sigma_z^2 = \sigma_y^2 \Longleftrightarrow 3^2 + 4^2 = 5^2$ relationship. In such case, we may have:

\begin{align*}
    p(x) &= \mathcal N(\mu_x = 0, \Sigma_x = 3^2) \\
    p(y) &= \mathcal N(\mu_y = 1, \Sigma_y = 4^2) \\
    p(z) &= \mathcal N(\mu_x+\mu_z,\Sigma_x+\Sigma_z) = \mathcal N(\mu_x = 1, \Sigma_x = 5^2)
\end{align*}

The plots will look like:

\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{{fig/e2.3}.png}
\end{figure}

We may tell these two plots are close resemblance of one and another by inspecting the shape of graph and the parameter of the fitted curve (showed as title of each plot).\newline

Please refer to \ilc{code/p2.py} for the code implementation.

\section*{E3. Dimensionality Reduction and PCA}

\section*{E4. Gaussian Mixture Models }

\end{document}

### 2021-04-21 A4/code/e3.py
⚙️ Add code for A3 E3.

@@ -1,0 +1,61 @@
+# https://www.datacamp.com/community/tutorials/principal-component-analysis-in-python
+# https://www.askpython.com/python/examples/principal-component-analysis
+# https://pub.towardsai.net/principal-component-analysis-pca-with-python-examples-tutorial-67a917bae9aa
+# https://jakevdp.github.io/PythonDataScienceHandbook/05.09-principal-component-analysis.html
+import pandas as pd
+import numpy as np
+import seaborn as sb
+import matplotlib.pyplot as plt
+from sklearn.decomposition import PCA
+
+
+def my_PCA(input , n):
+    standardized_input = input - np.mean(input , axis = 0)
+    cov_matrix = np.cov(standardized_input , rowvar = False)
+    eigen_values, eigen_vectors = np.linalg.eigh(cov_matrix)
+
+    eigen_value_sorted_index = np.argsort(eigen_values)[::-1]
+    sorted_eigenvalue = eigen_values[eigen_value_sorted_index]
+    sorted_eigenvectors = eigen_vectors[:,eigen_value_sorted_index]
+
+    first_n_eigenvectors = sorted_eigenvectors[:, 0 : n]
+    reduced_input = np.dot(first_n_eigenvectors.transpose(), standardized_input.transpose()).transpose()
+
+    return reduced_input
+
+
+
+
+
+# df = pd.read_csv('./A4/code/e3_breast_cancer.csv')
+# df = df.iloc[:, : -1]
+# label_column_name = 'diagnosis'
+# label = df.pop(label_column_name)
+
+
+df = pd.read_csv('./A4/code/e3_glass.csv')
+label_column_name = 'Type'
+label = df.pop(label_column_name)
+
+print(df.shape)
+
+
+reduce_to_n = 2
+column_names = ['PC'+str(i) for i in range(1, reduce_to_n+1)]
+reduced_df = pd.DataFrame(my_PCA(df , reduce_to_n) , columns = column_names)
+
+reduced_label_df = pd.concat([reduced_df , pd.DataFrame(label)] , axis = 1)
+
+# plt.figure(figsize = (10, 10))
+# sb.scatterplot(data = reduced_label_df, x = 'PC1', y = 'PC2', hue = label_column_name, s = 50, legend='full', palette= 'icefire')
+
+pca = PCA().fit(df)
+plt.plot(np.cumsum(pca.explained_variance_ratio_), color='b', label = 'cumulative')
[…]

### 2021-04-21 A4/article.tex
🐛 💄 Fix typos in E1.

@@
 
-Similarily, we may find out $\overline{\Sigma}$ as (note that $cov(x, y) = cov(y, x)$ for scalar):
+Similarily, we may find out $\overline{\Sigma}$ as (note that $cov(x, y) = cov(y, x)$):
 
@@
 
-As know from \textbf{Exercise 2.1} that $p(y) = \mathcal N(\mu_x+\mu_z,\Sigma_x+\Sigma_z)$ and $p(x) = \mathcal N(\mu_x, \Sigma_x)$, we have $p(y \mid x) = \mathcal N(\overline{\mu}, \overline{\mu})$, where:
+As know from \textbf{Exercise 2.1} that $p(y) = \mathcal N(\mu_x+\mu_z,\Sigma_x+\Sigma_z)$ and $p(x) = \mathcal N(\mu_x, \Sigma_x)$, we have $p(y \mid x) = \mathcal N(\overline{\mu}, \overline{\Sigma})$, where:
 
 \begin{align*}
-    \overline{\mu} &= \mu_y + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) = \mu_x \mu_z + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) \\
+    \overline{\mu} &= \mu_y + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) = \mu_x + \mu_z + \Sigma_{yx} \Sigma_{xx}^{-1} (x - \mu_x) \\
     \overline{\Sigma} &=  \Sigma_{yy} - \Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy} = \Sigma_{x} + \Sigma_{z} -\Sigma_{yx} \Sigma_{xx}^{-1} \Sigma_{xy}

### 2021-04-21 A4/code/e3.py
🔨 ⚙️ Update code for E3 printouts.

@@ -21,7 +21,7 @@ def my_PCA(input , n):
     first_n_eigenvectors = sorted_eigenvectors[:, 0 : n]
     reduced_input = np.dot(first_n_eigenvectors.transpose(), standardized_input.transpose()).transpose()
 
-    return reduced_input
+    return reduced_input, sorted_eigenvalue, sorted_eigenvectors
 
 
 
@@ -37,25 +37,36 @@ df = pd.read_csv('./A4/code/e3_glass.csv')
 label_column_name = 'Type'
 label = df.pop(label_column_name)
 
+print(df.head())
 print(df.shape)
 
 
 reduce_to_n = 2
 column_names = ['PC'+str(i) for i in range(1, reduce_to_n+1)]
-reduced_df = pd.DataFrame(my_PCA(df , reduce_to_n) , columns = column_names)
+reduced_df, sorted_eigenvalue, sorted_eigenvectors = my_PCA(df , reduce_to_n)
+reduced_df = pd.DataFrame(reduced_df, columns = column_names)
+reduced_label_df= pd.concat([reduced_df , pd.DataFrame(label)] , axis = 1)
 
-reduced_label_df = pd.concat([reduced_df , pd.DataFrame(label)] , axis = 1)
+# # 3.2.
+# print(sorted_eigenvalue)
+# print(sorted_eigenvectors)
+# print(f'<v_1, v_2> = {np.dot(sorted_eigenvectors[0], sorted_eigenvectors[1])}')
+# print(f'<v_2, v_3> = {np.dot(sorted_eigenvectors[1], sorted_eigenvectors[2])}')
+# print(f'<v_1, v_3> = {np.dot(sorted_eigenvectors[0], sorted_eigenvectors[2])}')
 
-# plt.figure(figsize = (10, 10))
-# sb.scatterplot(data = reduced_label_df, x = 'PC1', y = 'PC2', hue = label_column_name, s = 50, legend='full', palette= 'icefire')
 
-pca = PCA().fit(df)
-plt.plot(np.cumsum(pca.explained_variance_ratio_), color='b', label = 'cumulative')
-plt.plot(pca.explained_variance_ratio_, color='r', label = 'individual')
-plt.xlabel('number of components')
-plt.xticks([i for i in range(df.shape[1])])
-# plt.xticks([i for i in range(df.shape[1], 2)])
-plt.legend()
-plt.ylabel('explained variance');
+
+plt.figure(figsize = (10, 10))
[…]

### 2021-04-21 A4/article.tex
✨ Add answers to E3.

@@
 
+\subsection*{3.1.}
+
+The dataset I used in this exercise is the UCI Machine Learning \textit{Glass Classification} dataset I obtained from \href{https://www.kaggle.com/uciml/glass}{Kaggle}.
+
+\begin{lstlisting}
+        RI     Na    Mg    Al     Si     K    Ca   Ba   Fe  Type
+0  1.52101  13.64  4.49  1.10  71.78  0.06  8.75  0.0  0.0     1
+1  1.51761  13.89  3.60  1.36  72.73  0.48  7.83  0.0  0.0     1
+2  1.51618  13.53  3.55  1.54  72.99  0.39  7.78  0.0  0.0     1
+3  1.51766  13.21  3.69  1.29  72.61  0.57  8.22  0.0  0.0     1
+4  1.51742  13.27  3.62  1.24  73.08  0.55  8.07  0.0  0.0     1
+...
+(214, 10)
+\end{lstlisting}
+
+As showned above, it is a dataset of 214 sample, each with 9 attributes, and each sample has a label that belong to 7 of the potential glass types.
+The attributes is rather straightforward, with \ilc{RI} being refractive index, and all the others are  chemistry elements measured by weight percent in corresponding oxide (e.g.
+\ilc{Na} is Sodium, \ilc{Mg} is Magnesium, etc).
+The output labels are something among the lines of \ilc{container}, \ilc{tableware}, and \ilc{headlamps}.
+
+I opted to use this dataset as this is one of the few lightweight datasets with continues variables that I may find, and it has more than 6 dimensions.
+It probably doesn't make much sense to illustrate the dataset further, but the Kaggle page I linked above has some distribution plots for each element -- if you are interested.
+One thing to note about this dataset is, due to its lightweight, the dataset is really unbalanced as it has \ilc{70} samples on \ilc{Type 1}, \ilc{76} samples on \ilc{Type 1}, but only \ilc{17, 13, 9, 29} samples on \ilc{Type 3, 5, 6, 7} respectively (yes, \ilc{Type 4: vehicle} is entirely missing).
+
+\subsection*{3.2.}
+
+As we can't nicely plot 9-dimensional data. The first three eigenvalues and their corresponding eigenvectors are:
+
+\begin{lstlisting}
+3.00200916e+00 1.65917340e+00 6.79576475e-01
+
+[-9.28126899e-04  1.52290883e-03 -1.37689385e-03  3.10643441e-04
+-7.12950233e-04 -1.82174928e-03 -3.32594524e-04 -4.12235755e-03 9.99986948e-01]
+[-1.72248332e-02 -3.98797552e-01 -6.54934730e-01 -3.46599960e-01
+3.98381798e-01  1.55680962e-02 -3.76900981e-02 -3.62242832e-01 -1.39622177e-03]
+[ 7.23534913e-01  5.43050989e-01 -1.31198879e-01 -9.86931157e-02
+-7.68490459e-02  4.77602532e-02 -7.49534298e-02 -3.75274748e-01 -1.84522571e-03]
+\end{lstlisting}
+
+First we may confirm that the eigenvalues are indeed sorted in decreasing order, suggesting an eigenvalue with smaller index implies greater squared sum of variances and therefore accounts for higher percentage of variation.
+
+We may also confirm that these eigenvectors are indeed 9-dimensional as expected, and they are orthognal to eachother as the inner products between them yield 0:
+
+\begin{lstlisting}
+<v_1, v_2> = 4.336808689942018e-19
+<v_2, v_3> = -5.551115123125783e-17
+<v_1, v_3> = -4.336808689942018e-19
+\end{lstlisting}
+(There are not exactly zero due to datatype accuracy problem, but with at least $10^{-17}$ they are close enough to zero.)
+
+\subsection*{3.3.}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\linewidth]{{fig/e3.3}.png}
+\end{figure}
+
+The \textcolor{red}{red} line represents the percentage of variation that particular principle component accounts for.
+In this case, as we are ploting with respect to the decreasing order of eigenvalues, each principle component accountss for less variation than the previous one.
+Reversely, the \textcolor{blue}{blue}, whic represents the cumulative variance, is increasing as we add in more principle components -- and when we have all principle components (which is the dimension of the attributes or number of samples, in this case it is the former) considered, the cummulative explained variance reaches 1.
+
+\subsection*{3.4.}
+
+By reducing the dataset to 2 dimensions (\ilc{PC1, PC2}) -- meaning that these two eigenvectors has the two largest eigenvalues, suggesting the range of data projection are more distinguishiable on these two eigenvectors and therefore they contain most information -- we may have the following plot.
+Note the data were standardized with respect to their mean in this plot and throughout the PCA process, so that we won't have one catagory of data dominating all the others:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{{fig/e3.4_1}.png}
+\end{figure}
+
+I guess we may say the first two principle components perform pretty well on \ilc{Type 7} glass, but not much the others, especially between \ilc{Type 1, 2, 3}, a zoom in plot will look like this:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\linewidth]{{fig/e3.4_2}.png}
+\end{figure}
+
+This is something rather expected as we have 7 output labels (in fact, 6 output labels as we have no \ilc{Type 4}), but only 2 dimensions to project the clustering.
+The performance is further limited due to the unbalanced nature of having significantly more \ilc{Type 1} and \ilc{Type 2} samples, granting \ilc{Type 1\&2} glasses to display more features that are potentially similar to other types of glasses (just like doing multiple testing).
+
 \section*{E4. Gaussian Mixture Models }

### 2021-04-21 A4/code/e3.py
🔨 Adjust E3.4 plot size.

@@ -56,7 +56,7 @@ reduced_label_df= pd.concat([reduced_df , pd.DataFrame(label)] , axis = 1)
 
 
 
-plt.figure(figsize = (10, 10))
+plt.figure()
 sb.scatterplot(data = reduced_label_df, x = 'PC1', y = 'PC2', hue = label_column_name, s = 50, legend='full', palette= 'icefire')
 
 # 3.3.

### 2021-04-21 A4/code/exploration.py
⚙️ Add code for exploration.

@@ -1,0 +1,44 @@
+# https://towardsdatascience.com/visualising-high-dimensional-datasets-using-pca-and-t-sne-in-python-8ef87e7915b
+
+import numpy as np
+import pandas as pd
+import seaborn as sb
+import matplotlib.pyplot as plt
+# from sklearn.datasets import fetch_mldata
+from sklearn.decomposition import PCA
+from sklearn.manifold import TSNE
+
+
+# df = pd.read_csv('./A4/code/e3_breast_cancer.csv')
+# df = df.iloc[:, : -1]
+# label_column_name = 'diagnosis'
+# label = df.pop(label_column_name)
+
+df = pd.read_csv('./A4/code/e3_glass.csv')
+label_column_name = 'Type'
+label = df.pop(label_column_name)
+
+print(df.head())
+print(df.shape)
+
+tsne = TSNE(n_components=2, verbose=1, perplexity=100, n_iter=300)
+tsne_results = tsne.fit_transform(df)
+reduce_to_n = 2
+column_names = ['T-SNE '+str(i) for i in range(1, reduce_to_n+1)]
+reduced_df = pd.DataFrame(tsne_results, columns = column_names)
+reduced_label_df = pd.concat([reduced_df , pd.DataFrame(label)] , axis = 1)
+
+# df['T-SNE 1'] = tsne_results[:,0]
+# df['T-SNE 2'] = tsne_results[:,1]
+
+
+plt.figure()
+sb.scatterplot(
+    x="T-SNE 1", y="T-SNE 2",
+    hue=label_column_name,
+    palette='icefire',
+    data=reduced_label_df,
+    legend="full",
+    alpha=0.9
+)
+plt.show()

### 2021-04-21 A4/code/e1.py
⚙️ Add code for E1.

@@ -1,0 +1,100 @@
+# Ref. https://matplotlib.org/3.2.2/gallery/statistics/confidence_ellipse.html#sphx-glr-gallery-statistics-confidence-ellipse-py
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+import matplotlib.transforms as transforms
+
+def confidence_ellipse(x, y, ax, n_std=3.0, facecolor='none', **kwargs):
+    """
+    Create a plot of the covariance confidence ellipse of *x* and *y*.
+
+    Parameters
+    ----------
+    x, y : array-like, shape (n, )
+        Input data.
+
+    ax : matplotlib.axes.Axes
+        The axes object to draw the ellipse into.
+
+    n_std : float
+        The number of standard deviations to determine the ellipse's radiuses.
+
+    Returns
+    -------
+    matplotlib.patches.Ellipse
+
+    Other parameters
+    ----------------
+    kwargs : `~matplotlib.patches.Patch` properties
+    """
+    if x.size != y.size:
+        raise ValueError("x and y must be the same size")
+
+    cov = np.cov(x, y)
+    pearson = cov[0, 1]/np.sqrt(cov[0, 0] * cov[1, 1])
+    # Using a special case to obtain the eigenvalues of this
+    # two-dimensionl dataset.
+    ell_radius_x = np.sqrt(1 + pearson)
+    ell_radius_y = np.sqrt(1 - pearson)
+    ellipse = Ellipse((0, 0),
+        width=ell_radius_x * 2,
+        height=ell_radius_y * 2,
+        facecolor=facecolor,
+        **kwargs)
+
+    # Calculating the stdandard deviation of x from
+    # the squareroot of the variance and multiplying
+    # with the given number of standard deviations.
+    scale_x = np.sqrt(cov[0, 0]) * n_std
+    mean_x = np.mean(x)
+
+    # calculating the stdandard deviation of y ...
+    scale_y = np.sqrt(cov[1, 1]) * n_std
+    mean_y = np.mean(y)
+
+    transf = transforms.Affine2D() \
+        .rotate_deg(45) \
+        .scale(scale_x, scale_y) \
+        .translate(mean_x, mean_y)
+
+    ellipse.set_transform(transf + ax.transData)
[…]

### 2021-04-21 A4/article.tex
🐛 💄 Fix typos in E1.
✨ Add answers to E1.

@@
 \section*{E1. Multivariate Gaussians}
+
+Please refer to \ilc{A4/code/e1.py} for the code implmentations of below plots.
+
+\subsection*{1.1.}
+
+For 2D multivariate normal we have $\Sigma = \qbm{\sigma_{xx} & \sigma_{xy} \\ \sigma_{yx} & \sigma_{yy}}$. Note:
+
+\begin{itemize}
+    \item For $x$ and $y$ to be uncorrelated, we have $\sigma_{xy} = \sigma_{yx} = 0$. For demo, we have $\mathcal N(\mu = \qbm{0 \\ 0}, \Sigma = \qbm{1 & 0 \\ 0 & 1})$.
+    \item For $x$ and $y$ to be uncorrelated, we have $\sigma_{xy} = \sigma_{yx} > 0$. For demo, we have $\mathcal N(\mu = \qbm{0.5 \\ 0.5}, \Sigma = \qbm{1 & 0.9 \\ 0.9 & 1})$.
+    \item For $x$ and $y$ to be anti-correlated, we have $\sigma_{xy} = \sigma_{yx} < 0$. For demo, we have $\mathcal N(\mu = \qbm{0.7 \\ 0.7}, \Sigma = \qbm{1 & -0.9 \\ -0.9 & 1})$.
+\end{itemize}
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e1.1_1}.png}
+  \caption*{Uncorrelated}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e1.1_2}.png}
+    \caption*{Correlated}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e1.1_3}.png}
+    \caption*{Anti-correlated}
+\endminipage
+\end{figure}
+
+\subsection*{1.2.}
+
+Assuming the same $\mu$ and $\Sigma$ setup in \textbf{Exercise 1.1}, we have the $1-, 2-, 3-\sigma$ countour plots being (note eigenvectors are scaled \ilc{4x} for visual purposes):
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e1.2_1}.png}
+  \caption*{Uncorrelated}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e1.2_2}.png}
+    \caption*{Correlated}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e1.2_3}.png}
+    \caption*{Anti-correlated}
+\endminipage
+\end{figure}
+
 \section*{E2. Linear Gaussian Models}
@@
 
-We further know that there must be $p(y \mid x) = \mathcal N(\overline{\mu}, \overline{\mu})$, where:
+We further know that there must be $p(y \mid x) = \mathcal N(\overline{\mu}, \overline{\Sigma})$, where:
 
@@
 
+\section*{Exploration}
+
+% https://www.youtube.com/watch?v=NEaUSP4YerM
+% https://medium.com/@violante.andre/an-introduction-to-t-sne-with-python-example-47e6ae7dc58f
+
 \end{document}

### 2021-04-22 A4/article.tex
✨ Add writeup for Exploration.

@@
         E(z | x) &= E(z) = E(y + Ax) = \mu_y + A \mu_x = \mu_y -\Sigma_{yx} \Sigma_{xx}^{-1} \mu_x \\
-        E(z | x) &= E(y + Ax \mid x) = E(y \mid x) + A E(x \mid x)
+        E(z | x) &= E(y + Ax \mid x) = E(y \mid x) + A E(x \mid x) \\
     \end{cases} \\
@@
 
-\section*{E3. Dimensionality Reduction and PCA}
+\section*{E3. dimensionality Reduction and PCA}
 
@@
 
-\section*{Exploration}
+\section*{Exploration: t-SNE}
+
+\subsection*{Motivation}
 
@@
 
+There are many dimensionality reduction algorithms, as we have learned in this class, PCA is a dimensionality reduction algorithm that works fairly well on linear datasets and it is known to be used in combination with distance-based clustering algorithms.
+Intuitively, this makes PCA unapplicable of capturing dataset with non-linear structure -- such as a manifold ``Swiss roll'' we showed in class.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.4\linewidth]{{fig/exp_swiss_roll}.png}
+\end{figure}
+
+Well there are certainly PCA-based algorithms that are designed to handle this problem, e.g., the various avalivable and kernel PCAs, the SNEs and t-SNE in particular attracted my attention as it is a dimensionality reduction algorithm that doesn't wotk with distance-based clustering algorithms, yet yields promising results (\href{https://stats.stackexchange.com/questions/340175/why-is-t-sne-not-used-as-a-dimensionality-reduction-technique-for-clustering-or}{ref.}):
+
+\begin{figure}[H]
+\minipage{0.5\textwidth}
+  \includegraphics[width=\linewidth]{{fig/exp_pca_minist}.png}
+\endminipage\hfill
+\minipage{0.5\textwidth}
+    \includegraphics[width=\linewidth]{{fig/exp_tsne_minist}.png}
+\endminipage
+\end{figure}
+
+Clearly the t-SNE algorithms shows better clusters in comparison to PCA.
+So in this exploration I will give a brief walk-through on how t-SNE works, run it on the same dataset I used in \textbf{Exercise 3} and interpret the result, then discuss about the advantages between t-SNE and PCA in terms of determinability, locality, usability in combination with clustering algorithms, computational costs, etc.\newline
+
+\subsection*{Mechanism of t-SNE}
+
+As we already familiar from learning PCA, any direct projection of high-dimensional datapoints to lower dimensions will likely result in a mishmash:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.4\linewidth]{{fig/exp_direct_pj}.png}
+\end{figure}
+
+Intuitively, we would like to move around these projected datapoints, so that there can be a distinguishiable distance between different labels.
+However, the question is how?
+Which point should we move?
+And in which direction?
+Assume we have the following raw projection of two dimensional data onto one line:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\linewidth]{{fig/exp_tsne_raw}.png}
+\end{figure}
+
+t-SNE runs in an iterative manner, as it will move one point at a time untill it reaches the stopping criteria (which is usually numbers of iterations).
+W.L.O.G., let's check out how will t-SNE move the \textcolor{red}{leftmost red} point:
+
+\begin{enumerate}
+    \item First t-SNE will datapoints which is similar to this \textcolor{red}{leftmost red}, which are other \textcolor{red}{red} datapoints.
+    \item So naturally, we should move this \textcolor{red}{leftmost red} datapoint closer to other \textcolor{red}{red} datapoints -- which is to its right.
+    \item Similarily, since this \textcolor{red}{leftmost red} is far away from the \textcolor{orange}{orange} and \textcolor{blue}{blue} datapoints in its original space, it will like to move away from them -- which is to its left.
+\end{enumerate}
+
+So we have the following ``tractions'':
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\linewidth]{{fig/exp_tsne_demo}.png}
+\end{figure}
+
+Let's assume this time the traction from the \textcolor{red}{red} points are stronger, so the \textcolor{red}{leftmost red} moves to its right for a bit.
+Then we do this evaluation to each and every points, over and over, untill the stopping criteria is reached, and we may have some nice clusters on this one dimensionl line.\newline
+
+So how does t-SNE determine the similarity between datapoints?
+It does it, again, in an iterative manner.
+t-SNE will locate a point first (in this case, it will be the \textbf{black}{black} point from the \textbf{blue}{blue} cluster), and model its relative distances to all other datapoints under a normal distribution curve.
+So datapoints with same or similar lable, will have a higher probablity score, as it is closer to the center of the normal curve; yet datapoints that is vast unlike to eachother will have a lower probablity score, as it will be at the tails of normal curve.
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/exp_tsne_step1.1}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/exp_tsne_step1.2}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/exp_tsne_step1.3}.png}
+\endminipage
+\end{figure}
+
+t-SNE will then ``scale'' these probablity scores to unsure the total probably will be added to 1.
+In other words, assume datapoint $A$ has relative probablity scores of $p(A_x), p(A_y), \dots p(A_n)$ to datapoint $X, Y, \dots, N$, we will have $p(A_i') = \frac{p(A_i)}{\sum\limits_{j}^{n} p(A_j)}$.
+Following this procedure, t-SNE will obtain the relative normalized probablity score between every two datapoints, and thus fill this similarity matrix (note that it is common that $p(A_b') \neq p(B_a')$ as datapoints $A, B$ are under two different normal distributions.
+So t-SNE simple take the average out of these two) -- and the end result will be the rightmost matrix above; where \textcolor{red}{red entries} represent higher similarity.\newline
+
+Now we repeat such matrix filling operation with respect to the one-dimensional line, as we will fill in another matrix where the noralized probablity score will be determined by the relative distance between one point to another on this one-dimensional line.
+However, instead of modeling its with normal distribution, it will use t-distribution (and therefore, the ``t'' of ``t-SNE'') -- this is because t-distribution curve has a lower center and higher tails in comparison to normal distribution, thus encourges more seperation of datapoints.
+
+\begin{figure}[H]
+\minipage{0.7\textwidth}
+  \includegraphics[width=\linewidth]{{fig/exp_tsne_step2.1}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/exp_tsne_step2.2}.png}
+\endminipage
+\end{figure}
+
+With the two similarity matrices, t-SNE may then try to ``convert'' the second matrix more like the first one -- and the steps of this conversion will determine the moving direction of each datapoints.
+
+\subsection*{t-SNE on UCI ML Glass Classification Dataset}
+
+I have implemented t-SNE on the UCI ML Glass Classification Dataset I used in \textbf{Exercise 3} and got the following result:
+
+\begin{figure}[H]
+\minipage{0.5\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e3.4_1}.png}
+  \caption*{PCA}
+\endminipage\hfill
+\minipage{0.5\textwidth}
+    \includegraphics[width=\linewidth]{{fig/exp_glass}.png}
+    \caption*{t-SNE}
+\endminipage
+\end{figure}
+
+Although we don't have a clean clustered result like t-SNE on \ilc{MINIST} did.
+It is observable that we have better clustering result as the \ilc{Type 5, Type 6, Type 7} glasses are better gathered.
+Although we still can't effectively distinguish \ilc{Type 1} from \ilc{Type 2} glasses, they are certainly more scattered.
+
+(please refer to \ilc{A4/code/exploration} for code implementation.)
+
+\subsection*{t-SNE vs PCA}
+
+Although t-SNE offers better clustering visualization on many popular tasks, it also has its pros and cons.
+The main pros, are that t-SNE can certainly able to capture manifold structure, and it is better on visualization due to the procedure of t-SNE encourages different datapoints to ``move away from eachother,'' where PCA only cares about the maximum of sum of squared variance.\newline
+
+t-SNE also has many cons, and one of the most significant one is probably its incompability with distanced-based clustering algorithms.
+As we are directly moving the datapoints around in a lower-dimensional space, these datapoints lose their original distance relationship and thus become incapable of being a direct feed of distance-based clustering algorithms.
+Also on this note, t-SNE is not capable of learning a high-dimensional space to a lower-dimensions, but rather just lower the dimensions of known data -- so if an unseen data joins, it can't directly project it onto a lower-dimensionl space.\newline
+
+Further t-SNE is a non-determistic algorithm as it has a turnable paramater \textit{perplexity}, which is a loose balancer between the local and global aspect of given data, so you might have to run multiple trials to achive the optimal result.
+This is especially a problem as t-SNE is a lot more computational intense than PCA due to 1) it has to calculate the similarity scores that is proporsional to the combinations of datapoints, but not simply project all datapoints on to a line; 2) it has to do it iteraively over and over.
+
+The effect of this parameter is complicate, but according to \href{https://distill.pub/2016/misread-tsne/}{many articles} it works best if it is a rough estimation of cluster size.
+This implies t-SNE is probably not so suitable to unbalanced dataset (like the UCI ML Glass Classification I used).
+PCA, on the other hand, is parameter-free, but I don't consider this to be such a major advantage as the countless kernel PCA algorithms has much more turnable parameters.\newline
+
+A more through discussion between t-SNE and PCA can be found \href{https://stats.stackexchange.com/questions/238538/are-there-cases-where-pca-is-more-suitable-than-t-sne}{here} and \href{https://stats.stackexchange.com/questions/263539/clustering-on-the-output-of-t-sne}{here}.
+
 \end{document}

### 2021-04-22 A4/article.tex
🔨 Update E3.4 with more explaination.

@@
 
+Note the magnitutde of the entries of each eigenvalues are similar, with the exception that the first eigenvector shows greater value on element \ilc{Fe}, the first eigenvector shows lesser value on element \ilc{Mg}, and the third eigenvector shows greater value on \ilc{RI} (the reflective index) -- suggesting that the classification of glass samples might be postively correlated to \ilc{Fe} and \ilc{RI}, and negatively correlated to \ilc{Mg}.
+In terms of co-vary between eigenvectors, it is hard to tell if there is much consistency.
+
 \subsection*{3.3.}

### 2021-04-22 A4/code/e4.py
⚙️ Add code for E4.

@@ -1,0 +1,196 @@
+# Ref. https://towardsdatascience.com/gaussian-mixture-modelling-gmm-833c88587c7f
+# https://matplotlib.org/3.2.2/gallery/statistics/confidence_ellipse.html#sphx-glr-gallery-statistics-confidence-ellipse-py
+# http://ethen8181.github.io/machine-learning/clustering/GMM/GMM.html#Parameter-Estimation
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.stats import multivariate_normal
+from matplotlib.patches import Ellipse
+from sklearn import datasets
+import matplotlib.transforms as transforms
+
+np.random.seed(0)
+
+def confidence_ellipse(x, y, ax, n_std=3.0, mean=[], cov=[], facecolor='none', **kwargs):
+    """
+    Create a plot of the covariance confidence ellipse of *x* and *y*.
+
+    Parameters
+    ----------
+    x, y : array-like, shape (n, )
+        Input data.
+
+    ax : matplotlib.axes.Axes
+        The axes object to draw the ellipse into.
+
+    n_std : float
+        The number of standard deviations to determine the ellipse's radiuses.
+
+    Returns
+    -------
+    matplotlib.patches.Ellipse
+
+    Other parameters
+    ----------------
+    kwargs : `~matplotlib.patches.Patch` properties
+    """
+    if x.size != y.size:
+        raise ValueError("x and y must be the same size")
+
+    cov = np.cov(x, y) if len(cov) == 0 else cov
+    mean_x = np.mean(x) if len(mean) == 0 else mean[0]
+    mean_y = np.mean(y) if len(mean) == 0 else mean[1]
+
+    pearson = cov[0, 1]/np.sqrt(cov[0, 0] * cov[1, 1])
+    # Using a special case to obtain the eigenvalues of this
+    # two-dimensionl dataset.
+    ell_radius_x = np.sqrt(1 + pearson)
+    ell_radius_y = np.sqrt(1 - pearson)
+    ellipse = Ellipse((0, 0),
+        width=ell_radius_x * 2,
+        height=ell_radius_y * 2,
+        facecolor=facecolor,
+        **kwargs)
+
+    # Calculating the stdandard deviation of x from
+    # the squareroot of the variance and multiplying
+    # with the given number of standard deviations.
+    scale_x = np.sqrt(cov[0, 0]) * n_std
+
+    # calculating the stdandard deviation of y ...
[…]

### 2021-04-22 A4/article.tex
✨ Add answeres to E4.

@@
 
+Please refer to \ilc{A4/code/e4.py} for code implementation.
+Note the plotting is refered to \href{https://matplotlib.org/3.2.2/gallery/statistics/confidence_ellipse.html#sphx-glr-gallery-statistics-confidence-ellipse-py}{this}, and the GMM is refered to \href{http://ethen8181.github.io/machine-learning/clustering/GMM/GMM.html#Parameter-Estimation}{this} tutorials.
+
+\subsection*{4.1}
+
+\section{4.1}
+The EM algorithm iteratively performs the \ilc{expectation} step and \ilc{maximization} step -- thus ``EM'' -- until the log-likelihood (the indicator of convergence) falls into an acceptable \ilc{thlr}.
+
+\subsection*{Expectation Step}
+
+The expectation step intends to solve the inference problem: which Gaussian distribution generated each of the data point?
+By applying Baye's rule, the conditional probability of giving a data point $z_i = m$ given $z_i \in N_k$ can be represented as:
+\begin{align*}
+    \gamma_k = p(z_i=m|N_k) &= \frac{p(z_i=k)p(N_k|z_i=m)}{p(N_k)} \\
+    &= \frac{p(z_i=k)p(N_k|z_i=m)}{\sum_{j=1}^Kp(z_i=j)p(N_k|z_i=j)} \\
+    &= \frac{\pi_k\mathcal N(N_k|\mu_k, \Sigma_k)}{\sum_{j=1}^K \mathcal N(N_k|\mu_j, \Sigma_j)}
+\end{align*}
+Based on the parameters $\gamma_k$, $\pi_k$, $\mu_k$, and $\Sigma_k$, we are able to compute the log-likelihood for the purpose of optimization:
+\begin{align*}
+    \sum_k\sum_i \gamma_k^{(i)}log(\pi_k) + \sum_k\sum_i \gamma_{k}^{(i)}log(\mathcal N(N_k^{(i)|\mu_k, \Sigma_k}))
+\end{align*}
+The stopping criteria for the iteration is when the change of likelihood falls into a tolerance range that we predefined. We also need to define the
+maximum number of iterations in order to prevent it runs into infinite loop.
+
+\subsection*{Maximization Step}
+
+Since the change of the membership of data points will also affect the distributions, we can derive a closed form updates for all the parameters (the means $\mu$, the responsibility $\pi$, and the covariance matrix $\Sigma$).
+\begin{align*}
+    \mu_k &= \frac{1}{N_k}\sum_{n=1}^{N}\gamma_k^{(n)}x^{(n)} \\
+    \Sigma_{k} &= \frac{1}{N_k}\sum_{n=1}^{N}(x^{n}-\mu_{k})(x^{n}-\mu_{k})^T \\
+    \pi_k &= \frac{N_k}{N}
+\end{align*}
+, where $N_k=\Sigma_{n=1}^{N}\gamma_{k}^{(n)}$. We will perform E-step and M-step interchangeably to optimize the classification of data points.
+
+\subsection*{4.2}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\linewidth]{{fig/e4.2}.png}
+\end{figure}
+
+\subsection*{4.3}
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e4.3_0}.jpg}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e4.3_1}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.3_2}.png}
+\endminipage
+\end{figure}
+
+We can tell that all species from the Iris dataset (on two attributes) are nicely clustered within their sigma contours.
+
+\subsection*{4.4}
+
+\subsubsection*{K = 1}
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e4.4_k1_0}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.4_k1_5}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.4_k1_10}.png}
+\endminipage
+\end{figure}
+
+Where the $\mu, \Sigma, \pi$ for $K = 1$ are respectively:
+\begin{lstlisting}
+mean: [[ 3.48778309 70.89705882]]
+cov: [[[  1.29793889  13.92641885]
+  [ 13.92641885 184.14381488]]]
+pi: [1.]
+
+mean: [[ 3.48778309 70.89705882]]
+cov: [[[  1.29793889  13.92641885]
+  [ 13.92641885 184.14381488]]]
+pi: [1.]
+
+mean: [[ 3.48778309 70.89705882]]
+cov: [[[  1.29793889  13.92641885]
+  [ 13.92641885 184.14381488]]]
+pi: [1.]
+\end{lstlisting}
+
+\subsubsection*{K = 2}
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e4.4_k2_0}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.4_k2_5}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.4_k2_10}.png}
+\endminipage
+\end{figure}
+
+Where the $\mu, \Sigma, \pi$ for $K = 2$ are respectively:
+\begin{lstlisting}
+
+mean: [[ 4.32816932 81.43143106]
+[ 2.71366246 61.19333632]]
+cov: [[[  0.14492959   0.45361294]
+  [  0.45361294  28.18429193]]
+
+[[  1.11020876  10.67014897]
+  [ 10.67014897 131.42085961]]]
+pi: [0.47947807 0.52052193]
+
+mean: [[ 4.32816932 81.43143106]
+[ 2.71366246 61.19333632]]
+cov: [[[  0.14492959   0.45361294]
+  [  0.45361294  28.18429193]]
+[[  1.11020876  10.67014897]
+  [ 10.67014897 131.42085961]]]
+pi: [0.47947807 0.52052193]
+
+mean: [[ 4.32816932 81.43143106]
+[ 2.71366246 61.19333632]]
+cov: [[[  0.14492959   0.45361294]
+  [  0.45361294  28.18429193]]
+
+[[  1.11020876  10.67014897]
+  [ 10.67014897 131.42085961]]]
+pi: [0.47947807 0.52052193]
+\end{lstlisting}
+
+\subsubsection*{K = 3}
+
+\begin{figure}[H]
+\minipage{0.3\textwidth}
+  \includegraphics[width=\linewidth]{{fig/e4.4_k3_0}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.4_k3_5}.png}
+\endminipage\hfill
+\minipage{0.3\textwidth}
+    \includegraphics[width=\linewidth]{{fig/e4.4_k3_10}.png}
+\endminipage
+\end{figure}
+Where the $\mu, \Sigma, \pi$ for $K = 3$ are respectively:
+\begin{lstlisting}
+
+mean: [[ 4.23203926 83.42947492]
+[ 2.08519156 55.01970556]
+[ 4.34563196 78.61168712]]
+cov: [[[ 0.1506054   0.52994426]
+  [ 0.52994426 25.80940221]]
+[[ 0.13317075  1.1482405 ]
+  [ 1.1482405  41.28312871]]
+[[ 0.1526638   1.03335337]
+  [ 1.03335337 30.27117787]]]
+pi: [0.20632691 0.36913676 0.42453633]
+
+mean: [[ 4.23203926 83.42947492]
+[ 2.08519156 55.01970556]
+[ 4.34563196 78.61168712]]
+cov: [[[ 0.1506054   0.52994426]
+  [ 0.52994426 25.80940221]]
+[[ 0.13317075  1.1482405 ]
+  [ 1.1482405  41.28312871]]
+[[ 0.1526638   1.03335337]
+  [ 1.03335337 30.27117787]]]
+pi: [0.20632691 0.36913676 0.42453633]
+
+mean: [[ 4.23203926 83.42947492]
+[ 2.08519156 55.01970556]
+[ 4.34563196 78.61168712]]
+cov: [[[ 0.1506054   0.52994426]
+  [ 0.52994426 25.80940221]]
+[[ 0.13317075  1.1482405 ]
+  [ 1.1482405  41.28312871]]
+[[ 0.1526638   1.03335337]
+  [ 1.03335337 30.27117787]]]
+pi: [0.20632691 0.36913676 0.42453633]
+\end{lstlisting}
+
 \section*{Exploration: t-SNE}

### 2021-04-22 A4/article.tex
🔨 Add reference for Exploration.

@@
 t-SNE runs in an iterative manner, as it will move one point at a time untill it reaches the stopping criteria (which is usually numbers of iterations).
-W.L.O.G., let's check out how will t-SNE move the \textcolor{red}{leftmost red} point:
+W.L.O.G., let's check out how will t-SNE move the \textcolor{red}{leftmost red} point (note the following plots are coming from \href{https://www.youtube.com/watch?v=NEaUSP4YerM}{StatQuest with Josh Starmer} with a little bit post-process for better representation in written media):
 
